### 2024-01-13 doc/DataLiteracyStudentProject/paper.tex
first step in writing the paper

%%%%%%%% ICML 2023 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

\usepackage{tikz}
% Corporate Design of the University of Tübingen
% Primary Colors
\definecolor{TUred}{RGB}{165,30,55}
\definecolor{TUgold}{RGB}{180,160,105}
\definecolor{TUdark}{RGB}{50,65,75}
\definecolor{TUgray}{RGB}{175,179,183}

% Secondary Colors
\definecolor{TUdarkblue}{RGB}{65,90,140}
\definecolor{TUblue}{RGB}{0,105,170}
\definecolor{TUlightblue}{RGB}{80,170,200}
\definecolor{TUlightgreen}{RGB}{130,185,160}
\definecolor{TUgreen}{RGB}{125,165,75}
\definecolor{TUdarkgreen}{RGB}{50,110,30}
\definecolor{TUocre}{RGB}{200,80,60}
\definecolor{TUviolet}{RGB}{175,110,150}
\definecolor{TUmauve}{RGB}{180,160,150}
\definecolor{TUbeige}{RGB}{215,180,105}
\definecolor{TUorange}{RGB}{210,150,0}
\definecolor{TUbrown}{RGB}{145,105,70}

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2023} with \usepackage[nohyperref]{icml2023} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

\usepackage[accepted]{icml2023}

% For theorems and such
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Project Report Template for Data Literacy 2023/24}

\begin{document}

\twocolumn[
\icmltitle{My Data Literacy Project\\ (Replace this with your Project Title)}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2023
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Abdullah Abdul-Latif}{equal,first}
\icmlauthor{Lisa-Maria Fritsch}{equal,second}
\icmlauthor{Paul Kaifler}{equal,third}
\icmlauthor{Maximilian Schnitt}{equal,fourth}
\end{icmlauthorlist}

% fill in your matrikelnummer, email address, degree, for each group member
\icmlaffiliation{first}{Matrikelnummer 12345678, first.last@student.uni-tuebingen.de, MSc Machine Learning}
\icmlaffiliation{second}{Matrikelnummer 12345678, first.last@student.uni-tuebingen.de, MSc Computer Science}
\icmlaffiliation{third}{Matrikelnummer 12345678, first.last@student.uni-tuebingen.de, MSc Media Informatics}
\icmlaffiliation{fourth}{Matrikelnummer 6040570, maximilian.schnitt@student.uni-tuebingen.de, BSc Computer Science}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

%Put your abstract here. Abstracts typically start with a sentence motivating why the subject is interesting. Then mention the data, methodology or methods you are working with, and describe results. 
\begin{abstract}
As often mentioned in the news and in the daily life, people suggest that a higher 
\end{abstract}
%Motivate the problem, situation or topic you decided to work on. 
% Describe why it matters (is it of societal, economic, scientific value?). 
% Outline the rest of the paper (use references, e.g.~to \Cref{sec:methods}: 
% What kind of data you are working with, how you analyse it, and what kind of conclusion you reached. 
% The point of the introduction is to make the reader want to read the rest of the paper.

\section{Introduction}\label{sec:intro}
As often mentioned in the news and in the daily life, people suggest that a higher expected salary influences the job aspirations of the youth. 
We are curious about the underlying truth of thgis stetement. We have three Datasets, one Datset shows the Expected Salaries for various industrial sectors,
the second one shows the amount of students in the STEM sector in Tübingen from the Wintersemester 2005/2006 to the Wintersemester 2023/2024. 
The third dataset contains information about the amount of students, who finish high school in the certain year.
We did a feature analysis of the three datasets to get a better understanding of the data but the main focus of our analysis is on predicting 
the amount of students in the STEM sector for the next years. We used different approaches like a Random Forest approach with SARIMA
and a Multivariate Autoregression Model to investigate which approach works best for our purpose.

% In this section, describe \emph{what you did}. Roughly speaking, 
% explain what data you worked with, how or from where it was collected, it's structure and size.
%  Explain your analysis, and any specific choices you made in it. Depending on the nature of your project,
% you may focus more or less on certain aspects. If you collected data yourself, explain the collection process in detail.
% If you downloaded data from the net, show an exploratory analysis that builds
% intuition for the data, and shows that you know the data well. 
% If you are doing a custom analysis, explain how it works and why it is the right choice.
% If you are using a standard tool, it may still help to briefly outline it. Cite relevant works. 
% You can use the \verb|\citep| and \verb|\citet| commands for this purpose \citep{mackay2003information}.
\section{Data and Methods}\label{sec:methods}
We used three different pre collected datasets, our University data was collected by the Statistics Department from the University of Tuebingen. 
They used the immatriculation data from the students to create this dataset, which leads to
a quite accurate data which is necessary for the usage in our models.
\textcolor{red}{// HERE WE NEED MORE INFORMATION ABOUT THE DATA FROM THE OTHER DATASETS//}\\
The first step was to do some basic feature engineering on the datasets. We used various plots like line graphs, bar charts and pie charts
to get a better understanding of the data and detect trends. 
After we finished the basic feauture engineering we started the main analysis of the Vector Autoregressive Model (VAR). 
To get the most of this model, we wrote a parser, which parses text data generated from a PDF. This parser basically uses regular expressions to extract the amount of students
per years divided into the semesters and created a new dataset with the usage of pandas.\\
We've chosen the vector autoregressive model, because we have multivariate data and want to get a deeper understanding, if one of the datasets affects the others, 
so that we are able to predict one time series with another. A vector autoregressive model doesn't predict a single time series analyses, 
but multiple. Each time series is a linear combination which consists of n lags and an error term. A lag is the data from the past which we use to estimate
a new data point. We define the amount of lags in the hyperparameter p which defines the maxamount of lags which are used in the linear combination.
This linear combination also uses an error term, because we can't guarantee that our model predicts the correct output.\\
\textcolor{red}{One crucial step in the usage of those AR models, is that we have to ensure that our data is stationary. 
Data is stationary if those three attributes are fulfilled
\begin{enumerate}
    \item The mean $\mu$ is constant
    \item The standard deviation is constant
    \item There is no seasonality in the data
\end{enumerate}
// CHECK IF OUR DATA FULFILLS THOSE THREE ATTRIBUTES ///\\
After we ensured this, we did a PCAF to find statistically significant lags which we could use to find an optimal hyperparameter p of our vector autoregressive model.
A PCAF only represents the direct influence of the past datapoint $S_{n-i}$ for $i \in \{0,1,...,n\}$  to $S_n$. This gives us a graphical visualization with an error band of non statistical significant
lags in // GIVE A RANGE //.\\
After we used this PCAF we tried the different amount of lags, which looked like they have a significant input on the current datapoint $S_n$.
With the optimal lag we did a Granger Causality Test to get an understanding if one timeseries really influences another. The Granger Causality Test, does a t-Test and F-Test
checks if we can predict $S_i$ with $i \in \{2,3\}$ the time series $S_j$ $i\neq j$. The Granger Causality Test internally injects lags from the time series $S_J$ 
into the linear combination of $S_i$, then it does a t-Test and an F-test. As soon as one of those terms is verified by both tests,
then the test result is true.
}

% This is the template for a figure from the original ICML submission pack. In lecture 10 we will discuss plotting in detail.
% Refer to this lecture on how to include figures in this text.
% 
% \begin{figure}[ht]
% \vskip 0.2in
% \begin{center}
% \centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
% \caption{Historical locations and number of accepted papers for International
% Machine Learning Conferences (ICML 1993 -- ICML 2008) and International
% Workshops on Machine Learning (ML 1988 -- ML 1992). At the time this figure was
% produced, the number of accepted papers for ICML 2008 was unknown and instead
% estimated.}
% \label{icml-historical}
% \end{center}
% \vskip -0.2in
% \end{figure}

\section{Results}\label{sec:results}

In this section outline your results. At this point, you are just stating the outcome of your analysis. 
You can highlight important aspects (``we observe a significantly higher value of $x$ over $y$''), but leave interpretation and opinion to the next section. This section absoultely \emph{has} to include at least two figures.

\section{Discussion \& Conclusion}\label{sec:conclusion}

Use this section to briefly summarize the entire text. Highlight limitations and problems, but also make clear statements where they are possible and supported by the analysis.

\section*{Contribution Statement}

Explain here, in one sentence per person, what each group member contributed. For example, you could write: Max Mustermann collected and prepared data. Gabi Musterfrau and John Doe performed the data analysis. Jane Doe produced visualizations. All authors will jointly wrote the text of the report. Note that you, as a group, a collectively responsible for the report. Your contributions should be roughly equal in amount and difficulty.

\section*{Notes}

Your entire report has a \textbf{hard page limit of 4 pages} excluding references. (I.e. any pages beyond page 4 must only contain references). Appendices are \emph{not} possible. But you can put additional material, like interactive visualizations or videos, on a githunb repo (use \href{https://github.com/pnkraemer/tueplots}{links} in your pdf to refer to them). Each report has to contain \textbf{at least three plots or visualizations}, and \textbf{cite at least two references}. More details about how to prepare the report, inclucing how to produce plots, cite correctly, and how to ideally structure your github repo, will be discussed in the lecture, where a rubric for the evaluation will also be provided.

\bibliography{bibliography}
\bibliographystyle{icml2023}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
% Modified again in 2023 by Sivan Sabato and Jonathan Scarlett.
% Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2024-01-14 src/constants.py
update data path

@@ -4,7 +4,7 @@ import re
 
 # paths
 PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
-DATA_DIR = os.path.join(PROJECT_DIR, "data")
+DATA_DIR = os.path.join(PROJECT_DIR, "dat")
 GRADUATES_PATH = os.path.join(DATA_DIR, "graduates_germany.csv")
 SALLARY_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
 STUDENTS_PATH = os.path.join(DATA_DIR, "student_data_per_subject.xlsx")

### 2024-01-14 exp/RandomForestRegressor.py
formatting

@@ -3,8 +3,8 @@ from sklearn.model_selection import train_test_split
 from sklearn.ensemble import RandomForestRegressor
 from sklearn.metrics import mean_absolute_error
 
-import salaries 
-import students 
+import salaries
+import students
 import graduates
 
 import matplotlib.pyplot as plt
@@ -58,7 +58,6 @@ target_std = students_data['Students'].std()
 print("Standard Deviation of Target Variable:", target_std)
 
 
-
 # For forecasting future student numbers, you would predict using new salary and graduates data:
 future_salaries = salaries.GetSalaryDataForecast()
 future_graduates = graduates.GetGraduatesDataForecast()
@@ -87,4 +86,3 @@ ax2.set_ylabel('Salaries')
 ax2.legend(loc='upper left')
 
 plt.show()
-

### 2024-01-14 exp/graduates.py

@@ -8,7 +8,8 @@ from statsmodels.tsa.stattools import adfuller
 from statsmodels.graphics.tsaplots import plot_pacf
 from statsmodels.graphics.tsaplots import plot_acf
 
-import os 
+import os
+
 
 def GetGraduatesData():
 
@@ -23,9 +24,9 @@ def GetGraduatesData():
     PLOT_DIR = os.path.join(PROJECT_DIR, 'plots')
     """
 
-    #file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/graduates_germany.csv'
+    # file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/graduates_germany.csv'
     file_path = GRADUATES_PATH
-    data = pd.read_csv(file_path, encoding= "ISO-8859-1",sep=";", decimal=".", skiprows=4, skipfooter=3, engine="python")
+    data = pd.read_csv(file_path, encoding="ISO-8859-1", sep=";", decimal=".", skiprows=4, skipfooter=3, engine="python")
 
     # Now manually assign the first column as 'year' and the second as 'gender'
     data.columns = ['gender', 'year'] + ['state_' + str(i) for i in range(1, len(data.columns) - 1)]
@@ -37,7 +38,7 @@ def GetGraduatesData():
     data_with_years = data[data['year'].str.contains(r'20\d{2}/\d{2}', regex=True)]
 
     # Define the range of years we are interested in
-    desired_years = [str(year) + '/' + str(year+1)[-2:] for year in range(2006, 2022)]
+    desired_years = [str(year) + '/' + str(year + 1)[-2:] for year in range(2006, 2022)]
 
     # Now filter the DataFrame to only include the desired years
     data_years_filtered = data_with_years[data_with_years['year'].isin(desired_years)]
@@ -90,7 +91,7 @@ def GetGraduatesData():
     graduates.columns = columns_list
 
     # Display the final data with total male graduates per year, excluding the year with 0 graduates
-    #print(graduates)
+    # print(graduates)
 
     return graduates
 
@@ -117,7 +118,6 @@ def GetGraduatesDataForecast():
     plt.show()
     """
 
-
     """
     # Fit an ARIMA(1,0,0) model
[…]

### 2024-01-14 exp/salaries.py

@@ -8,7 +8,8 @@ from statsmodels.tsa.stattools import adfuller
 from statsmodels.graphics.tsaplots import plot_pacf
 from statsmodels.graphics.tsaplots import plot_acf
 
-import os 
+import os
+
 
 def GetSalaryData():
 
@@ -16,9 +17,9 @@ def GetSalaryData():
     DATA_DIR = os.path.join(PROJECT_DIR, "dat")
     GRADUATES_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
 
-    #file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/sallary_per_sector.csv'
+    # file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/sallary_per_sector.csv'
     file_path = GRADUATES_PATH
-    data = pd.read_csv(file_path, encoding= "ISO-8859-1",sep=";", decimal=".", skiprows=5, skipfooter=9, index_col=0, engine="python")
+    data = pd.read_csv(file_path, encoding="ISO-8859-1", sep=";", decimal=".", skiprows=5, skipfooter=9, index_col=0, engine="python")
 
     # Get the list of columns in the DataFrame
     columns_list = data.columns.tolist()
@@ -51,12 +52,18 @@ def GetSalaryData():
 
     # Preprocess the 'quarter' column to standard format
     # Assuming the original format is something like '1. Quartal', '2. Quartal', etc.
-    new_data_with_five_columns_male['quarter'] = new_data_with_five_columns_male['quarter'].str.extract('(\d)').astype(str)
-    new_data_with_five_columns_female['quarter'] = new_data_with_five_columns_female['quarter'].str.extract('(\d)').astype(str)
+    new_data_with_five_columns_male['quarter'] = new_data_with_five_columns_male['quarter'].str.extract('(\\d)').astype(str)
+    new_data_with_five_columns_female['quarter'] = new_data_with_five_columns_female['quarter'].str.extract('(\\d)').astype(str)
 
     # Now create the 'date' column
-    new_data_with_five_columns_male['date'] = pd.to_datetime(new_data_with_five_columns_male['year'].astype(str) + 'Q' + new_data_with_five_columns_male['quarter'])
[…]

### 2024-01-14 exp/students.py

@@ -9,18 +9,19 @@ from statsmodels.tsa.stattools import adfuller
 from statsmodels.graphics.tsaplots import plot_pacf
 from statsmodels.graphics.tsaplots import plot_acf
 
-import os 
+import os
+
 
 def GetStudentsData():
-    
+
     PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
     DATA_DIR = os.path.join(PROJECT_DIR, "dat")
     GRADUATES_PATH = os.path.join(DATA_DIR, "student_data_per_subject.csv")
 
-    #file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/student_data_per_subject.csv'
+    # file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/student_data_per_subject.csv'
     file_path = GRADUATES_PATH
 
-    students = pd.read_csv(file_path, encoding= "ISO-8859-1",sep=";", decimal=".", index_col=0, skiprows=0, skipfooter=0, engine="python")
+    students = pd.read_csv(file_path, encoding="ISO-8859-1", sep=";", decimal=".", index_col=0, skiprows=0, skipfooter=0, engine="python")
 
     # Get the list of columns in the DataFrame
     columns_list = students.columns.tolist()
@@ -58,10 +59,10 @@ def GetStudentsData():
     for i, start_index in enumerate(academic_year_indexes):
         # Add the academic year column name
         renamed_columns.append(columns_list[start_index])
-        
+
         # Determine the end index for this academic year's range of columns
         end_index = academic_year_indexes[i + 1] if i + 1 < len(academic_year_indexes) else len(columns_list)
-        
+
         # Generate the new Fachsemester names for the columns in this range
         for j in range(1, end_index - start_index):
             if j <= 12:
@@ -83,13 +84,12 @@ def GetStudentsData():
             column_series = pd.Series(students[col].squeeze())
             students[col] = pd.to_numeric(column_series, errors='coerce')
 
-
     for i, index in enumerate(academic_year_indexes):
         # Determine the range of columns for this academic year's Fachsemester
[…]

### 2024-01-14 src/constants.py

@@ -63,4 +63,3 @@ YEARS = [
     "SoSe 2023",
     "WiSe 2023/2024",
 ]
-

### 2024-01-14 src/dataLoading.py
Better relative imports

@@ -1,7 +1,7 @@
 """Utilities to load data into a pandas DataFrame."""
 
 import pandas as pd
-from constants import GRADUATES_PATH, STUDENTS_PATH, SALLARY_PATH, INFLATION_PATH
+from .constants import GRADUATES_PATH, STUDENTS_PATH, SALLARY_PATH, INFLATION_PATH
 
 
 def getStudents() -> pd.DataFrame:

### 2024-01-14 src/dataLoading.py
reafactor regression.ipynb

@@ -1,7 +1,8 @@
 """Utilities to load data into a pandas DataFrame."""
 
 import pandas as pd
-from .constants import GRADUATES_PATH, STUDENTS_PATH, SALLARY_PATH, INFLATION_PATH
+import numpy as np
+from .constants import GRADUATES_PATH, STUDENTS_PATH, SALLARY_PATH, INFLATION_PATH, YEARS, SEMESTERS
 
 
 def getStudents() -> pd.DataFrame:
@@ -61,3 +62,71 @@ def getInflation() -> pd.DataFrame:
         skipfooter=3,
         engine="python")
     return inflation
+
+
+def getAllCourses() -> list:
+    """Get all courses from student data."""
+
+    students = getStudents()
+    return students.index.levels[0].tolist()
+
+
+def getTotalStudentsFor(courses: list, years: list[str]) -> np.array:
+    students = getStudents()
+
+    studentsInCourseHF = np.empty((len(courses), 0))
+
+    for year in years:
+        data_for_year = students.loc[(courses, 'HF'), (year, SEMESTERS)].to_numpy(dtype=int).sum(axis=1)
+        data_for_year = data_for_year.reshape((len(courses), 1))
+        studentsInCourseHF = np.hstack((studentsInCourseHF, data_for_year))
+
+    studentsInCourseNF = np.empty((len(courses), 0))
+
+    for year in years:
+        data_for_year = students.loc[(courses, 'NF'), (year, SEMESTERS)].to_numpy(dtype=int).sum(axis=1)
+        data_for_year = data_for_year.reshape((len(courses), 1))
+        studentsInCourseNF = np.hstack((studentsInCourseNF, data_for_year))
+
+    return studentsInCourseHF.sum(axis=0) + studentsInCourseNF.sum(axis=0)
+
+
+def getBruttoSallary(sector) -> np.array:
+    sallaries = getSallaries()
+    SALLARY_YEARS = sallaries.index.levels[2]
+    QUARTALS = ['1. Quartal', '2. Quartal', '3. Quartal', '4. Quartal']
+
+    bruttoSallary = sallaries.loc[(sector[0], sector[1], SALLARY_YEARS, QUARTALS), ('Insgesamt', 'Insgesamt',
+                                                                                    'Durchschnittliche Bruttomonatsverdienste', 'EUR')].to_numpy(dtype=int)
+    bruttoSallary = bruttoSallary.reshape(-1, 2)
[…]

### 2024-01-14 exp/RandomForestRegressor.py
Add main.py and LaTeX compilation functionality
- Implemented main.py script that automates the generation of figures and compilation of LaTeX documents into PDF.
- The script utilizes functions from RandomForestRegressor.py to generate required figures.
- Added LaTeX PDF compilation command that converts .tex files to .pdf using pdflatex, enabling direct production of paper drafts from LaTeX source files.
- This update is intended to streamline the report generation process for the predicting_students project.

@@ -2,89 +2,104 @@ import pandas as pd
 from sklearn.model_selection import train_test_split
 from sklearn.ensemble import RandomForestRegressor
 from sklearn.metrics import mean_absolute_error
+import os 
 
-import salaries 
-import students 
-import graduates
+from . import salaries
+from . import students
+from . import graduates
 
 import matplotlib.pyplot as plt
 
+def RandomForestRegressorPrediction():
 
-# Load your datasets
-salaries_data = salaries.GetSalaryData()
-graduates_data = graduates.GetGraduatesData()
-students_data = students.GetStudentsData()
+    # Load your datasets
+    salaries_data = salaries.GetSalaryData()
+    graduates_data = graduates.GetGraduatesData()
+    students_data = students.GetStudentsData()
 
-""""
-print(salaries_data.shape)
-print(graduates_data.shape)
-print(students_data.shape)
-"""
+    """" Check shape of dataframes
+    print(salaries_data.shape)
+    print(graduates_data.shape)
+    print(students_data.shape)
+    """
 
-# Combine datasets into a single DataFrame (assuming they are all aligned and sorted by date)
-combined_data = pd.DataFrame({
-    'salaries': salaries_data['Salary'],
-    'graduates': graduates_data['Graduates'],
-    'students': students_data['Students']
-})
+    # Combine datasets into a single DataFrame (assuming they are all aligned and sorted by date)
+    combined_data = pd.DataFrame({
+        'salaries': salaries_data['Salary'],
+        'graduates': graduates_data['Graduates'],
+        'students': students_data['Students']
+    })
 
-# Define features and target variable
-X = combined_data[['salaries', 'graduates']]  # Features
-y = combined_data['students']  # Target
+    # Define features and target variable
+    X = combined_data[['salaries', 'graduates']]  # Features
+    y = combined_data['students']  # Target
 
-# Split data into training and testing set
-X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.2, random_state=42)
+    # Split data into training and testing set
+    X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.2, random_state=42)
 
-# Initialize the Random Forest Regressor
-model = RandomForestRegressor(n_estimators=100, random_state=42)
+    # Initialize the Random Forest Regressor
+    model = RandomForestRegressor(n_estimators=100, random_state=42)
 
-# Fit the model
[…]

### 2024-01-14 exp/graduates.py

@@ -146,7 +146,7 @@ def GetGraduatesDataForecast():
     forecast = model_fit.forecast(steps=8) # Forecasting next 5 periods
     """
 
-
+    """
     
     plt.figure(figsize=(10, 6))
     plt.plot(time_series_data['Graduates'], label='Historical')
@@ -155,6 +155,7 @@ def GetGraduatesDataForecast():
     plt.show()
 
     plt.close()
+    """
     
     return forecast
 

### 2024-01-14 exp/salaries.py

@@ -14,10 +14,10 @@ def GetSalaryData():
 
     PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
     DATA_DIR = os.path.join(PROJECT_DIR, "dat")
-    GRADUATES_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
+    SALARY_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
 
     #file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/sallary_per_sector.csv'
-    file_path = GRADUATES_PATH
+    file_path = SALARY_PATH
     data = pd.read_csv(file_path, encoding= "ISO-8859-1",sep=";", decimal=".", skiprows=5, skipfooter=9, index_col=0, engine="python")
 
     # Get the list of columns in the DataFrame

### 2024-01-14 exp/students.py

@@ -15,10 +15,10 @@ def GetStudentsData():
     
     PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
     DATA_DIR = os.path.join(PROJECT_DIR, "dat")
-    GRADUATES_PATH = os.path.join(DATA_DIR, "student_data_per_subject.csv")
+    STUDENTS_PATH = os.path.join(DATA_DIR, "student_data_per_subject.csv")
 
     #file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/student_data_per_subject.csv'
-    file_path = GRADUATES_PATH
+    file_path = STUDENTS_PATH
 
     students = pd.read_csv(file_path, encoding= "ISO-8859-1",sep=";", decimal=".", index_col=0, skiprows=0, skipfooter=0, engine="python")
 
@@ -193,7 +193,7 @@ def GetStudentsDataForecast():
     forecast = model_fit.forecast(steps=8) # Forecasting next 5 periods
     """
 
-    
+    """
     plt.figure(figsize=(10, 6))
     plt.plot(time_series_data['Students'], label='Historical')
     plt.plot(forecast, label='Forecast')
@@ -202,6 +202,7 @@ def GetStudentsDataForecast():
 
     plt.close()
     
+    """
 
     return forecast
 

### 2024-01-14 src/main.py

@@ -1,0 +1,26 @@
+import os
+import subprocess
+import matplotlib.pyplot as plt
+import sys
+
+# Add the parent directory to the Python path
+sys.path.append(os.path.dirname(os.path.dirname(os.path.abspath(__file__))))
+from exp import RandomForestRegressor as rfg
+
+
+# Generate figures using python scripts 
+def generate_figures():
+    rfg.RandomForestRegressorPrediction()
+
+# Compile LaTeX paper into PDF (if applicable)
+def compile_paper():
+    PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
+    paper_tex_path = os.path.join(PROJECT_DIR,"doc/DataLiteracyStudentProject/samplePaper.tex")
+    paper_pdf_path = os.path.join(PROJECT_DIR,"doc/DataLiteracyStudentProject/samplePaper.pdf")
+
+    subprocess.run(['pdflatex', '-output-directory', os.path.dirname(paper_pdf_path), paper_tex_path])
+
+if __name__ == "__main__":
+    # Generate and save figures
+    generate_figures()
+    compile_paper()

### 2024-01-15 src/constants.py
latex to pdf workflow

@@ -4,10 +4,16 @@ import re
 
 # paths
 PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
-DATA_DIR = os.path.join(PROJECT_DIR, "data")
+DATA_DIR = os.path.join(PROJECT_DIR, "dat")
 GRADUATES_PATH = os.path.join(DATA_DIR, "graduates_germany.csv")
 SALLARY_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
 STUDENTS_PATH = os.path.join(DATA_DIR, "student_data_per_subject.xlsx")
+EXP_DIR = os.path.join(PROJECT_DIR, "exp")
+
+# paths for the paper
+PAPER_DIR = os.path.join(PROJECT_DIR, "doc/DataLiteracyStudentProject")
+TEX_PATH = os.path.join(PAPER_DIR, "paper.tex")
+FIG_DIR = os.path.join(PAPER_DIR, "fig")
 
 
 # REGEX Patterns

### 2024-01-15 src/constants.py
create fig folder

@@ -14,6 +14,9 @@ EXP_DIR = os.path.join(PROJECT_DIR, "exp")
 PAPER_DIR = os.path.join(PROJECT_DIR, "doc/DataLiteracyStudentProject")
 TEX_PATH = os.path.join(PAPER_DIR, "paper.tex")
 FIG_DIR = os.path.join(PAPER_DIR, "fig")
+# Create the folder if not present
+if not os.path.exists(FIG_DIR):
+    os.makedirs(FIG_DIR)
 
 
 # REGEX Patterns

### 2024-01-15 exp/graduates.py
Rm ARIMA model

@@ -1,162 +1,0 @@
-import pandas as pd
-from statsmodels.tsa.arima.model import ARIMA
-from matplotlib import pyplot as plt
-import seaborn as sns
-from statsmodels.tsa.statespace.sarimax import SARIMAX
-
-from statsmodels.tsa.stattools import adfuller
-from statsmodels.graphics.tsaplots import plot_pacf
-from statsmodels.graphics.tsaplots import plot_acf
-
-import os 
-
-def GetGraduatesData():
-
-    PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
-    DATA_DIR = os.path.join(PROJECT_DIR, "dat")
-    GRADUATES_PATH = os.path.join(DATA_DIR, "graduates_germany.csv")
-
-    """
-    SALLARY_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
-    GENDER_DATA_PATH = os.path.join(DATA_DIR, "university_gender_divided_stem.xlsx")
-    STUDENTS_PATH = os.path.join(DATA_DIR, "student_data_per_subject.xlsx")
-    PLOT_DIR = os.path.join(PROJECT_DIR, 'plots')
-    """
-
-    #file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/graduates_germany.csv'
-    file_path = GRADUATES_PATH
-    data = pd.read_csv(file_path, encoding= "ISO-8859-1",sep=";", decimal=".", skiprows=4, skipfooter=3, engine="python")
-
-    # Now manually assign the first column as 'year' and the second as 'gender'
-    data.columns = ['gender', 'year'] + ['state_' + str(i) for i in range(1, len(data.columns) - 1)]
-
-    # Convert the 'year' to string for consistency
-    data['year'] = data['year'].astype(str)
-
-    # Filter out rows that do not contain the desired year format (e.g., '2006/07')
-    data_with_years = data[data['year'].str.contains(r'20\d{2}/\d{2}', regex=True)]
-
-    # Define the range of years we are interested in
-    desired_years = [str(year) + '/' + str(year+1)[-2:] for year in range(2006, 2022)]
-
-    # Now filter the DataFrame to only include the desired years
-    data_years_filtered = data_with_years[data_with_years['year'].isin(desired_years)]
-
[…]

### 2024-01-15 exp/salaries.py

@@ -1,163 +1,0 @@
-import pandas as pd
-from statsmodels.tsa.arima.model import ARIMA
-from matplotlib import pyplot as plt
-import seaborn as sns
-from statsmodels.tsa.statespace.sarimax import SARIMAX
-
-from statsmodels.tsa.stattools import adfuller
-from statsmodels.graphics.tsaplots import plot_pacf
-from statsmodels.graphics.tsaplots import plot_acf
-
-import os 
-
-def GetSalaryData():
-
-    PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
-    DATA_DIR = os.path.join(PROJECT_DIR, "dat")
-    GRADUATES_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
-
-    #file_path = '/Users/abdallahabdul-latif/Desktop/Universität Tübingen/5.Semester/Data Literacy/StudentProject/AnalysingStudentDevelopment/data/sallary_per_sector.csv'
-    file_path = GRADUATES_PATH
-    data = pd.read_csv(file_path, encoding= "ISO-8859-1",sep=";", decimal=".", skiprows=5, skipfooter=9, index_col=0, engine="python")
-
-    # Get the list of columns in the DataFrame
-    columns_list = data.columns.tolist()
-
-    # Let's say you want to rename the column at index 2
-    columns_list[0] = 'sector'
-    columns_list[1] = 'year'
-    columns_list[2] = 'quarter'
-    columns_list[3] = 'salary'
-    columns_list[6] = 'salary'
-
-    # Assign the modified list of column names back to the DataFrame
-    data.columns = columns_list
-
-    # The value you're looking for
-    specific_value = 'Forschung und Entwicklung'
-
-    # Create a new DataFrame with only the rows where 'sector' column has the specific value
-    filtered_data = data[data['sector'] == specific_value].copy()
-
-    # Creating a new DataFrame with only the first five columns of the filtered_data DataFrame
-    new_data_with_five_columns = filtered_data.iloc[:, :7].copy()
-
-    new_data_with_five_columns_male = new_data_with_five_columns.iloc[:, :4].copy()
-    new_data_with_five_columns_female = new_data_with_five_columns.iloc[:, [0, 1, 2, -1]]
-
-    # Preprocess year column
-    new_data_with_five_columns_male['year'] = new_data_with_five_columns_male['year'].astype(int)
[…]

### 2024-01-15 src/dataLoading.py
rm old notebook

@@ -2,7 +2,7 @@
 
 import pandas as pd
 import numpy as np
-from .constants import GRADUATES_PATH, STUDENTS_PATH, SALLARY_PATH, INFLATION_PATH, YEARS, SEMESTERS
+from .constants import GRADUATES_PATH, STUDENTS_PATH, SALLARY_PATH, INFLATION_PATH, SEMESTERS
 
 
 def getStudents() -> pd.DataFrame:
@@ -96,8 +96,7 @@ def getBruttoSallary(sector) -> np.array:
     SALLARY_YEARS = sallaries.index.levels[2]
     QUARTALS = ['1. Quartal', '2. Quartal', '3. Quartal', '4. Quartal']
 
-    bruttoSallary = sallaries.loc[(sector[0], sector[1], SALLARY_YEARS, QUARTALS), ('Insgesamt', 'Insgesamt',
-                                                                                    'Durchschnittliche Bruttomonatsverdienste', 'EUR')].to_numpy(dtype=int)
+    bruttoSallary = sallaries.loc[(sector[0], sector[1], SALLARY_YEARS, QUARTALS), ('Insgesamt', 'Insgesamt', 'Durchschnittliche Bruttomonatsverdienste', 'EUR')].to_numpy(dtype=int)  # noqa: E501
     bruttoSallary = bruttoSallary.reshape(-1, 2)
     return bruttoSallary.mean(axis=1)
 

### 2024-01-15 src/constants.py
modified new data

@@ -5,7 +5,7 @@ import re
 # paths
 PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
 DATA_DIR = os.path.join(PROJECT_DIR, "dat")
-GRADUATES_PATH = os.path.join(DATA_DIR, "graduates_germany.csv")
+GRADUATES_PATH = os.path.join(DATA_DIR, "graduates_BW.csv")
 SALLARY_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
 STUDENTS_PATH = os.path.join(DATA_DIR, "student_data_per_subject.xlsx")
 INFLATION_PATH = os.path.join(DATA_DIR, "inflation_germany.csv")

### 2024-01-15 src/dataLoading.py
use new dataset

@@ -19,17 +19,16 @@ def getStudents() -> pd.DataFrame:
 def getGraduates() -> pd.DataFrame:
     """Load graduates data from csv file into a pandas DataFrame."""
 
-    graduates = pd.read_csv(
-        GRADUATES_PATH,
-        encoding="ISO-8859-1",
-        sep=";",
-        decimal=",",
-        skiprows=4,
-        header=[1, 2, 3],
-        skipfooter=3,
-        index_col=[0, 1],
-        engine="python")
-    return graduates
+    graduatesInBW = pd.read_csv(GRADUATES_PATH,
+                                sep=";",
+                                thousands=".",
+                                skiprows=1,
+                                header=[0, 1],
+                                skipfooter=10,
+                                index_col=0,
+                                na_values=["."],
+                                engine="python")
+    return graduatesInBW
 
 
 def getSallaries() -> pd.DataFrame:
@@ -103,14 +102,13 @@ def getBruttoSallary(sector) -> np.array:
 
 def getGraduatesInBwFor(years: list) -> np.array:
     graduates = getGraduates()
-    graduatesInBW = graduates.loc[('Insgesamt', years), ('Baden-Württemberg', 'Absolventen und Abgänger', 'Anzahl')].to_numpy()
-    graduatesInBW = np.where(graduatesInBW == '-', np.nan, graduatesInBW)
+    graduatesInBW = graduates.loc[years, ('Abiturienten insg', 'Anzahl')].to_numpy()
     return graduatesInBW.astype(float)
 
 
 def getAllGraduatesYears() -> list[str]:
     graduates = getGraduates()
-    return graduates.index.levels[1].tolist()
+    return graduates.index.tolist()
 
 
 def getInflationAdjustedBruttoSallary(sector) -> np.array:

### 2024-01-16 doc/DataLiteracyStudentProject/paper.tex
updated paper

@@
 \usepackage{tikz}
+\usepackage{enumitem}
 % Corporate Design of the University of Tübingen
@@
 \begin{abstract}
-As often mentioned in the news and in the daily life, people suggest that a higher 
-\end{abstract}
+As often mentioned in the news and in the daily life, people suggest that a higher expected salary is a strong pull factor for students 
+choosing their subject. We were curious if this is particullarily true for the scientific based subjects, because a majority of them have a 
+pretty high expected salary.
+We used the students data from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics University of Tübingen},
+the high-school graduates data from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg}
+and also the salary data from 2005 to 2023 from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt}.
+We used a vector autoregressive model, to build a time series forecast and analysed the outcomes to find an optimal configuration for this type of model in our context.
+\textcolor{red}{result.}
 %Motivate the problem, situation or topic you decided to work on. 
@@
 \section{Introduction}\label{sec:intro}
-As often mentioned in the news and in the daily life, people suggest that a higher expected salary influences the job aspirations of the youth. 
-We are curious about the underlying truth of thgis stetement. We have three Datasets, one Datset shows the Expected Salaries for various industrial sectors,
-the second one shows the amount of students in the STEM sector in Tübingen from the Wintersemester 2005/2006 to the Wintersemester 2023/2024. 
-The third dataset contains information about the amount of students, who finish high school in the certain year.
-We did a feature analysis of the three datasets to get a better understanding of the data but the main focus of our analysis is on predicting 
-the amount of students in the STEM sector for the next years. We used different approaches like a Random Forest approach with SARIMA
-and a Multivariate Autoregression Model to investigate which approach works best for our purpose.
+In the public eye, there is a strong correlation between the expected salary and the subject enrollment of students. We are curious if this statement is true or just a biased
+perception of the society. Our main focus is on the University of Tübingen and mainly on the STEM (scientific engineering technical and mathematical) subjects, because those degrees 
+are likely to have a pretty high expected salary after graduating. Furthermore we needed data of high-school graduates in Baden-Würrtemberg, to find a possible corellation between
+the amount of graduates and the new enrollments in the stem subjects. We just focused on Baden-Württemberg, this introduces a bias to our prediction, but we assumed that the 
+majority of students which are going to study in Tübingen come from Baden-Württemberg. The third datatset we used is about the salaries from 2005 to 2023 in the main 
+industrial sectors in Germany. The main focus of this paper is on a analysis of a vetor autoregressive model, to predict the upcoming numbers of students in the next semesters. We
+used various techniques to clean our data like removing the inflation or making our time series stationary to ensure that the predictions don't detect patterns which are inferred by 
+some noise in the data. Nevertheless we also used basic data science techniques to get a better understanding of the dirstributions and trends over the past years.
 
@@
 \section{Data and Methods}\label{sec:methods}
-We used three different pre collected datasets, our University data was collected by the Statistics Department from the University of Tuebingen. 
-They used the immatriculation data from the students to create this dataset, which leads to
-a quite accurate data which is necessary for the usage in our models.
-\textcolor{red}{// HERE WE NEED MORE INFORMATION ABOUT THE DATA FROM THE OTHER DATASETS//}\\
-The first step was to do some basic feature engineering on the datasets. We used various plots like line graphs, bar charts and pie charts
-to get a better understanding of the data and detect trends. 
-After we finished the basic feauture engineering we started the main analysis of the Vector Autoregressive Model (VAR). 
-To get the most of this model, we wrote a parser, which parses text data generated from a PDF. This parser basically uses regular expressions to extract the amount of students
-per years divided into the semesters and created a new dataset with the usage of pandas.\\
-We've chosen the vector autoregressive model, because we have multivariate data and want to get a deeper understanding, if one of the datasets affects the others, 
-so that we are able to predict one time series with another. A vector autoregressive model doesn't predict a single time series analyses, 
-but multiple. Each time series is a linear combination which consists of n lags and an error term. A lag is the data from the past which we use to estimate
-a new data point. We define the amount of lags in the hyperparameter p which defines the maxamount of lags which are used in the linear combination.
-This linear combination also uses an error term, because we can't guarantee that our model predicts the correct output.\\
-\textcolor{red}{One crucial step in the usage of those AR models, is that we have to ensure that our data is stationary. 
-Data is stationary if those three attributes are fulfilled
-\begin{enumerate}
+In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of theUniversity of Tübingen}., which contains the
+total amount of students from the winter semester 2005/2006 to the winter semester 2023/2024. The data was collected from the enrollments of students, which registered 
+for a study in Tübingen.
+The second dataset is from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} which contains data about the high-school graduates from 1970 to 2023.
+Our third and last dataset is from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt} and contains the salaries of the large industrial sectors in Germany from 2005 to 2021.
+One main problem with the mapping between university subjects and actual jobs is that there exists no perfect one-to-one mapping. 
+We also removed the inflation from the salaries to get the real wage, because the inflation could 
+cause a false pattern detection in our autoregressive model. 
+To reduce the amount of bias introduced in this step, we also used the Inflation data from the Deutsches Statistisches Bundesamt over the years.\\
+
+\textcolor{red}{To get the most of this model, we wrote a parser, which parses text data generated from a PDF. This parser basically uses regular expressions to extract the amount of students
+per years divided into the semesters and created a new dataset with the usage of pandas.\\}
+
+Our first steps involved some feauture engineering on the three different datasets to get a better understanding of the data and also detect patterns in the data like trends and distributions over time.
+After we finished the basic feauture engineering step we started the main analysis of the Vector Autoregressive Model (VAR). 
+We've chosen the vector autoregressive model, because we have multivariate data and we assumes that there is a correlation between the given time series.
+A VAR model don't predict a single time series but one for each input series. Those modles have one hyperparameter \textcolor{red}{p} which define the amount 
+of previous data points which should be represented in the output series.
+Each prediction is a linear combination with the form  $S_n = C_{1} \cdot S_{n-1}+ C_{2} \dots+ C_i \cdot  S_{n-i} + \epsilon$ for
+$i \in \{1,...,|lags|\}$. In our case we have three incoming time series which also leads to three outgoing time series. Our outgoing prediction for the data point n based on lags 
+is represented as
+\begin{equation*}
+    \resizebox{\linewidth}{!}{%
+        $\begin{bmatrix}
+            S_{1,n}\\
+            S_{2,n}\\
+            S_{3,n}
+        \end{bmatrix} = 
+        \begin{bmatrix}
+            C_{1,1} & C_{1,2} & \dots &C_{1,n-1}\\
+            C_{2,1} & C_{2,2} &\dots & C_{2,n-1}\\
+            C_{3,1} & C_{3,2} &\dots &C_{3,n-1}\\
+        \end{bmatrix} + 
+        \begin{bmatrix}
+            S_{1,1} & S_{1,2} & \dots & S_{1,n-1}\\
+            S_{2,1} & S_{2,2} & \dots & S_{2,n-1}\\
+            S_{3,1} & S_{3,2} & \dots & S_{3,n-1}\\
+        \end{bmatrix}+
+        \begin{bmatrix}
+            \epsilon_{1}\\
+            \epsilon_{2}\\
+            \epsilon_{3}
+        \end{bmatrix}$%
+    }
+    \end{equation*}
+Each time series $S_n$ $\forall i \in \{1,2,3\}$ describes the n-th data point for the input time series i. The term $\epsilon$ is added to approximate an error, which is 
+induced by the model.\\
+We need to assure that our model is stationary, which means that the the stochastic process doesn't change over time. \textcolor{red}{WHY?}.
+Our data is stationary if we are able to assure this for all our time series data inputs:
+\begin{enumerate}[label=\Roman*]
     \item The mean $\mu$ is constant
@@
 \end{enumerate}
-// CHECK IF OUR DATA FULFILLS THOSE THREE ATTRIBUTES ///\\
-After we ensured this, we did a PCAF to find statistically significant lags which we could use to find an optimal hyperparameter p of our vector autoregressive model.
-A PCAF only represents the direct influence of the past datapoint $S_{n-i}$ for $i \in \{0,1,...,n\}$  to $S_n$.
-This gives us a graphical visualization with an error band of non statistical significant
-lags in // GIVE A RANGE //.\\
-After we used this PCAF we tried the different amount of lags, which looked like they have a significant input on the current datapoint $S_n$.
-With the optimal lag we did a Granger Causality Test to get an understanding if one timeseries really influences another. The Granger Causality Test, does a t-Test and F-Test
-checks if we can predict $S_i$ with $i \in \{2,3\}$ the time series $S_j$ $i\neq j$. The Granger Causality Test internally injects lags from the time series $S_J$ 
-into the linear combination of $S_i$, then it does a t-Test and an F-test. As soon as one of those terms is verified by both tests,
-then the test result is true.
-}
+
+After we ensured, that our data is stationary, we used a PACF plot to find lags which could be statistically significant. 
+The main reason why we choose PACF in contrast to ACF is that we only want to look at direct influences from $S_{n-1}$ to $S_{n}$ and not from $S_{n-3}$ to $S_n$.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth,height=\textheight, keepaspectratio]{fig/pacf_plot.pdf}
+    \caption{PACF Plot for our VAR Model}
+    \label{fig:enter-label}
+\end{figure}
+\textcolor{red}{// CONCLUSION FOR PACF // }\\
+
+After we found the optimal lag configuration of our model, we used a Granger Causality Test, to statistically prove that the salary time series really influences the student decision.
+The main idea of Granger Causality is that we are able to predict the data point $S_{i,n}$ directly with one or multiple other time serieses. The Granger Causality Test uses exactly this
+idea, by inserting data points $S*:=S_{j,m}$ for $j\in i:j\neq i$  and $m<n$. After we inserted a couple of terms we do a t-Test and an F-Test on this new $S*$ and as soon as both tests
+return statistical significance for the same S* we can assure that the three time series are Granger Causal.
 

### 2024-01-17 src/salaryExpHelpers.py
exp: modify sallry expectation

@@ -1,0 +1,35 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def plot(data, forecast, years) -> plt.Figure:
+    dataWithForecast = np.vstack((data, forecast))
+    yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
+
+    fig, ax1 = plt.subplots()
+    plt.title('University Data Over the Years')
+    plt.xticks(rotation='vertical')
+    plt.axvspan(len(data), len(dataWithForecast) - 1, alpha=0.5, color='lightgrey')
+
+    ax1.set_xlabel('Year')
+    ax1.set_ylabel('Total students and Graduates in BW', color='tab:blue')
+    ax1.bar(yearsWithForecast, dataWithForecast[:, 0], color='tab:blue', label='Total students')  # Changed to bar
+    ax1.tick_params(axis='y', labelcolor='tab:blue')
+
+    ax2 = ax1.twinx()  # instantiate a second axes that shares the same x-axis
+
+    color = 'tab:red'
+    ax2.set_ylabel('Average brutto sallary', color=color)  # we already handled the x-label with ax1
+    ax2.plot(yearsWithForecast, dataWithForecast[:, 1], color=color, label='Average brutto sallary')
+    ax2.tick_params(axis='y', labelcolor=color)
+
+    # Create one legend for both subplots
+    lines, labels = ax1.get_legend_handles_labels()
+    lines2, labels2 = ax2.get_legend_handles_labels()
+    ax2.legend(lines + lines2, labels + labels2, loc='upper left')
+
+
+def modifySalary(data, func) -> np.ndarray:
+    modified_data = np.copy(data)
+    modified_data[:, 1] = np.array([func(d) for d in data[:, 1]])
+    return modified_data

### 2024-01-17 src/salaryExpHelpers.py
small enhancement

@@ -30,6 +30,5 @@ def plot(data, forecast, years) -> plt.Figure:
 
 
 def modifySalary(data, func) -> np.ndarray:
-    modified_data = np.copy(data)
-    modified_data[:, 1] = np.array([func(d) for d in data[:, 1]])
-    return modified_data
+    data[:, 1] = np.array([func(d) for d in data[:, 1]])
+    return data

### 2024-01-18 src/constants.py
fix typo

@@ -6,7 +6,7 @@ import re
 PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
 DATA_DIR = os.path.join(PROJECT_DIR, "dat")
 GRADUATES_PATH = os.path.join(DATA_DIR, "graduates_BW.csv")
-SALLARY_PATH = os.path.join(DATA_DIR, "sallary_per_sector.csv")
+SALARY_PATH = os.path.join(DATA_DIR, "salary_per_sector.csv")
 STUDENTS_PATH = os.path.join(DATA_DIR, "student_data_per_subject.xlsx")
 INFLATION_PATH = os.path.join(DATA_DIR, "inflation_germany.csv")
 EXP_DIR = os.path.join(PROJECT_DIR, "exp")

### 2024-01-19 src/constants.py
change colors

@@ -1,6 +1,7 @@
 # Description: Constants used in the code
 import os
 import re
+from tueplots.constants.color import rgb
 
 # paths
 PROJECT_DIR = os.path.dirname(os.path.dirname(os.path.abspath(__file__)))
@@ -19,6 +20,9 @@ FIG_DIR = os.path.join(PAPER_DIR, "fig")
 if not os.path.exists(FIG_DIR):
     os.makedirs(FIG_DIR)
 
+# colors
+STUDENT_COLOR = rgb.tue_blue
+SALARY_COLOR = rgb.tue_red
 
 # REGEX Patterns
 DELIMETER_PATTERN = re.compile(r".*\s*Studienfachbelegung nach Abschlusszielen")

### 2024-01-19 src/constants.py
change blue color

@@ -21,7 +21,7 @@ if not os.path.exists(FIG_DIR):
     os.makedirs(FIG_DIR)
 
 # colors
-STUDENT_COLOR = rgb.tue_blue
+STUDENT_COLOR = rgb.tue_lightblue
 SALARY_COLOR = rgb.tue_red
 
 # REGEX Patterns

### 2024-01-19 src/constants.py
Better plot

@@ -21,8 +21,10 @@ if not os.path.exists(FIG_DIR):
     os.makedirs(FIG_DIR)
 
 # colors
-STUDENT_COLOR = rgb.tue_lightblue
-SALARY_COLOR = rgb.tue_red
+COL_STUDENT = rgb.tue_lightblue
+COL_STUDENT_PRED = rgb.tue_blue
+COL_SALARY = rgb.tue_red
+COL_SALARY_PRED = rgb.tue_violet
 
 # REGEX Patterns
 DELIMETER_PATTERN = re.compile(r".*\s*Studienfachbelegung nach Abschlusszielen")

### 2024-01-19 src/constants.py
OK plots

@@ -24,7 +24,7 @@ if not os.path.exists(FIG_DIR):
 COL_STUDENT = rgb.tue_lightblue
 COL_STUDENT_PRED = rgb.tue_blue
 COL_SALARY = rgb.tue_red
-COL_SALARY_PRED = rgb.tue_violet
+COL_SALARY_PRED = rgb.tue_orange
 
 # REGEX Patterns
 DELIMETER_PATTERN = re.compile(r".*\s*Studienfachbelegung nach Abschlusszielen")

### 2024-01-19 src/salaryExpHelpers.py
Adapt code to changes

@@ -1,32 +1,44 @@
 import matplotlib.pyplot as plt
 import numpy as np
+from src.constants import COL_STUDENT, COL_STUDENT_PRED, COL_SALARY, COL_SALARY_PRED
+from tueplots import bundles
 
 
 def plot(data, forecast, years) -> plt.Figure:
     dataWithForecast = np.vstack((data, forecast))
     yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
 
+    plt.rcParams.update(bundles.icml2022(column="full", nrows=1, ncols=1, usetex=False))
     fig, ax1 = plt.subplots()
     plt.title('University Data Over the Years')
-    plt.xticks(rotation='vertical')
-    plt.axvspan(len(data), len(dataWithForecast) - 1, alpha=0.5, color='lightgrey')
+    plt.xticks(rotation=30)
 
-    ax1.set_xlabel('Year')
-    ax1.set_ylabel('Total students and Graduates in BW', color='tab:blue')
-    ax1.bar(yearsWithForecast, dataWithForecast[:, 0], color='tab:blue', label='Total students')  # Changed to bar
-    ax1.tick_params(axis='y', labelcolor='tab:blue')
+    ax1.set_ylabel('Total students', color=COL_STUDENT)
+    # Plot main data
+    ax1.bar(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 0], color=COL_STUDENT, label='Enrolled students')
+
+    # Plot forecast with a different color
+    ax1.bar(yearsWithForecast[-len(forecast):], dataWithForecast[-len(forecast):, 0],
+            color=COL_STUDENT_PRED, label='Enrolled students forecast')
 
     ax2 = ax1.twinx()  # instantiate a second axes that shares the same x-axis
+    ax2.set_ylabel('Sallary', color=COL_SALARY)
 
-    color = 'tab:red'
-    ax2.set_ylabel('Average brutto sallary', color=color)  # we already handled the x-label with ax1
-    ax2.plot(yearsWithForecast, dataWithForecast[:, 1], color=color, label='Average brutto sallary')
-    ax2.tick_params(axis='y', labelcolor=color)
+    # Plot main data
+    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average brutto salary')
+
+    # Plot forecast with a different color
[…]

### 2024-01-20 download_conda.sh
Update README and better names

@@ -1,3 +1,0 @@
-#!/bin/bash
-# creates environment
-conda env create -f data_literacy.yml 

### 2024-01-20 install_env.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+# creates environment
+conda env create -f conda_env.yml 

### 2024-01-21 src/dataLoading.py
Apply data flattening to all notebooks

@@ -28,6 +28,13 @@ def getGraduates() -> pd.DataFrame:
                                 index_col=0,
                                 na_values=["."],
                                 engine="python")
+    # Flatten data anamoly due to shift from G9 to G8 in 2012
+    newVal = (graduatesInBW.loc[2011, ('Abiturienten insg', 'Anzahl')] + graduatesInBW.loc[2013, ('Abiturienten insg', 'Anzahl')]) / 2
+    diff = graduatesInBW.loc[2012, ('Abiturienten insg', 'Anzahl')] - newVal
+    graduatesInBW.loc[2012, ('Abiturienten insg', 'Anzahl')] = newVal + diff / 4
+    graduatesInBW.loc[2013, ('Abiturienten insg', 'Anzahl')] += diff / 4
+    graduatesInBW.loc[2014, ('Abiturienten insg', 'Anzahl')] += + diff / 4
+    graduatesInBW.loc[2015, ('Abiturienten insg', 'Anzahl')] += + diff / 4
     return graduatesInBW
 
 

### 2024-01-24 src/salaryExpHelpers.py
fix salary bug

@@ -2,6 +2,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 from src.constants import COL_STUDENT, COL_STUDENT_PRED, COL_SALARY, COL_SALARY_PRED
 from tueplots import bundles
+import copy
 
 
 def plot(data, forecast, years) -> plt.Figure:
@@ -42,5 +43,6 @@ def plot(data, forecast, years) -> plt.Figure:
 
 
 def modifySalary(data, func) -> np.ndarray:
-    data[:, 1] = np.array([func(d) for d in data[:, 1]])
-    return data
+    newData = copy.deepcopy(data)
+    newData[:, 1] = np.array([func(d) for d in newData[:, 1]])
+    return newData

### 2024-01-25 src/constants.py
added savefig functionality

@@ -11,7 +11,9 @@ SALARY_PATH = os.path.join(DATA_DIR, "salary_per_sector.csv")
 STUDENTS_PATH = os.path.join(DATA_DIR, "student_data_per_subject.xlsx")
 INFLATION_PATH = os.path.join(DATA_DIR, "inflation_germany.csv")
 EXP_DIR = os.path.join(PROJECT_DIR, "exp")
+DOC_PATH = os.path.join(PROJECT_DIR, "doc", "DataLiteracyStudentProject")
 
+FIG_DIR = os.path.join(DOC_PATH, "fig")
 # paths for the paper
 PAPER_DIR = os.path.join(PROJECT_DIR, "doc/DataLiteracyStudentProject")
 TEX_PATH = os.path.join(PAPER_DIR, "paper.tex")

### 2024-01-25 src/salaryExpHelpers.py

@@ -1,11 +1,14 @@
 import matplotlib.pyplot as plt
-import numpy as np
-from src.constants import COL_STUDENT, COL_STUDENT_PRED, COL_SALARY, COL_SALARY_PRED
-from tueplots import bundles
+
 import copy
+import os
+
+import numpy as np
+from src.constants import COL_STUDENT, COL_STUDENT_PRED, COL_SALARY, COL_SALARY_PRED, FIG_DIR
+from tueplots import bundles
 
 
-def plot(data, forecast, years) -> plt.Figure:
+def plot(data, forecast, years, name, lags) -> plt.Figure:
     dataWithForecast = np.vstack((data, forecast))
     yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
 
@@ -26,11 +29,11 @@ def plot(data, forecast, years) -> plt.Figure:
     ax2.set_ylabel('Sallary', color=COL_SALARY)
 
     # Plot main data
-    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average brutto salary')
+    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average salary after tax')
 
     # Plot forecast with a different color
     ax2.plot(yearsWithForecast[-len(forecast) - 1:], dataWithForecast[-len(forecast) - 1:, 1],
-             color=COL_SALARY_PRED, label='Average brutto salary forecast')
+             color=COL_SALARY_PRED, label='Average salary after tax forecast')
 
     # Create one legend for both subplots
     lines, labels = ax1.get_legend_handles_labels()
@@ -40,6 +43,7 @@ def plot(data, forecast, years) -> plt.Figure:
     # Get current ticks, keep only every second tick
     locs, labels = plt.xticks()
     plt.xticks(locs[::2], labels[::2])
+    fig.savefig(os.path.join(FIG_DIR, f"{name}_{lags}.pdf"))
 
 
 def modifySalary(data, func) -> np.ndarray:

### 2024-01-25 doc/DataLiteracyStudentProject/paper.tex
Include a new fig
Visualization of mapping
First implementation

@@
 \begin{icmlauthorlist}
-\icmlauthor{Abdullah Abdul-Latif}{equal,first}
+\icmlauthor{Abdallah Abdul-Latif}{equal,first}
 \icmlauthor{Lisa-Maria Fritsch}{equal,second}
@@
 % fill in your matrikelnummer, email address, degree, for each group member
-\icmlaffiliation{first}{Matrikelnummer 12345678, first.last@student.uni-tuebingen.de, MSc Machine Learning}
+\icmlaffiliation{first}{Matrikelnummer 5977981, abdallah.abdul-latif@student.uni-tuebingen.de, MSc Computer Science}
 \icmlaffiliation{second}{Matrikelnummer 12345678, first.last@student.uni-tuebingen.de, MSc Computer Science}
@@
 % The point of the introduction is to make the reader want to read the rest of the paper.
+\end{abstract}
 
@@
 The main reason why we choose PACF in contrast to ACF is that we only want to look at direct influences from $S_{n-1}$ to $S_{n}$ and not from $S_{n-3}$ to $S_n$.
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\linewidth,height=\textheight, keepaspectratio]{fig/pacf_plot.pdf}
+\begin{figure}
+    \includegraphics{fig/pacf_plot.pdf}
     \caption{PACF Plot for our VAR Model}
-    \label{fig:enter-label}
 \end{figure}
@@
 This section absoultely \emph{has} to include at least two figures.
+
+\begin{figure*}
+    \includegraphics{fig/params.pdf}
+    \caption{This is a figure caption. Write this, if we end up using that fig ;-)}
+\end{figure*}
 

### 2024-01-28 src/constants.py
Augmented data
Now we are using all sectors and courses which we preselected and mapped together (for exp_01)

@@ -80,3 +80,16 @@ YEARS = [
     "SoSe 2023",
     "WiSe 2023/2024",
 ]
+
+COURSES = ['"Biologie"', '"Chemie"', '"Geowissenschaften"', '"Informatik"', '"Mathematik"', '"Pharmazie"', '"Biochemie"', '"Physik"', '"Psychologie"']
+SECTORS = ['Herstellung von pharmazeutischen Erzeugnissen',
+    'Herstellung von chemischen Erzeugnissen',
+    'Forschung und Entwicklung',
+    'Gewinnung von Steinen und Erden, sonstiger Bergbau',
+    'Dienstleistungen der Informationstechnologie',
+    'H.v. DV-Geräten, elektron. u. opt. Erzeugnissen',
+    'Wirtschaftsprüfung und Steuerberatung, Buchführung',
+    'H.v. Bestrahlungs- u.a. elektromed. Geräten',
+    'Gesundheitswesen',
+]
+

### 2024-01-28 src/dataLoading.py

@@ -3,6 +3,7 @@
 import pandas as pd
 import numpy as np
 from .constants import GRADUATES_PATH, STUDENTS_PATH, SALARY_PATH, INFLATION_PATH, SEMESTERS
+import os 
 
 
 def getStudents() -> pd.DataFrame:
@@ -49,9 +50,9 @@ def getSalaries() -> pd.DataFrame:
         skiprows=5,
         header=[0, 1, 2, 3],
         skipfooter=4,
-        index_col=[0, 1, 2, 3],
+        index_col=[1, 2, 3],
         engine="python")
-    return salaries
+    return salaries.drop(salaries.columns[0], axis=1)
 
 
 def getInflation() -> pd.DataFrame:
@@ -98,13 +99,14 @@ def getTotalStudentsFor(courses: list, years: list[str]) -> np.array:
 
 
 def getBruttoSalary(sector) -> np.array:
-    salaries = getSalaries()
-    SALARY_YEARS = salaries.index.levels[2]
+    salaries = getSalaries()    
+    SALARY_YEARS = salaries.index.levels[1]
     QUARTALS = ['1. Quartal', '2. Quartal', '3. Quartal', '4. Quartal']
 
-    bruttoSalary = salaries.loc[(sector[0], sector[1], SALARY_YEARS, QUARTALS), ('Insgesamt', 'Insgesamt', 'Durchschnittliche Bruttomonatsverdienste', 'EUR')].to_numpy(dtype=int)  # noqa: E501
-    bruttoSalary = bruttoSalary.reshape(-1, 2)
-    return bruttoSalary.mean(axis=1)
+    bruttoSalaryNP = salaries.loc[(sector, SALARY_YEARS, QUARTALS), ('Insgesamt', 'Insgesamt', 'Durchschnittliche Bruttomonatsverdienste', 'EUR')].to_numpy(dtype=int) # noqa: E501
+    bruttoSalaryMeanForYears = bruttoSalaryNP.reshape(-1, 2).mean(axis=1)
+    bruttoSalary = bruttoSalaryMeanForYears.reshape(-1,30).mean(axis=0)
+    return bruttoSalary
 
 
 def getGraduatesInBwFor(years: list) -> np.array:
@@ -134,3 +136,34 @@ def getInflationAdjustedBruttoSalary(sector) -> np.array:
     for a, b in zip(bruttoSalary, bruttoInflationSalary):
         print(f'{a} -> {b}')
     return bruttoInflationSalary
[…]

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
changed methods part

@@
 \section{Data and Methods}\label{sec:methods}
-In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of theUniversity of Tübingen}., which contains the
+In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}., which contains the
 total amount of students from the winter semester 2005/2006 to the winter semester 2023/2024. The data was collected from the enrollments of students, which registered 
@@
 
-\textcolor{red}{To get the most of this model, we wrote a parser, which parses text data generated from a PDF. This parser basically uses regular expressions to extract the amount of students
-per years divided into the semesters and created a new dataset with the usage of pandas.\\}
-
-Our first steps involved some feauture engineering on the three different datasets to get a better understanding of the data and also detect patterns in the data like trends and distributions over time.
-After we finished the basic feauture engineering step we started the main analysis of the Vector Autoregressive Model (VAR). 
-We've chosen the vector autoregressive model, because we have multivariate data and we assumes that there is a correlation between the given time series.
-A VAR model don't predict a single time series but one for each input series. Those modles have one hyperparameter \textcolor{red}{p} which define the amount 
-of previous data points which should be represented in the output series.
-Each prediction is a linear combination with the form  $S_n = C_{1} \cdot S_{n-1}+ C_{2} \dots+ C_i \cdot  S_{n-i} + \epsilon$ for
-$i \in \{1,...,|lags|\}$. In our case we have three incoming time series which also leads to three outgoing time series. Our outgoing prediction for the data point n based on lags 
-is represented as
+Our main goal was to find a correlation between the salary, the graduates and the resulting amount of students enrolling a subject in the STEM field.
+For this problem we use a vector autoregressive model, which predicts a time series
+for each input time series. The model creates a linear combination for each prediction which could be expressed in a matrix-vector notation.
 \begin{equation*}
@@
             S_{2,n}\\
-            S_{3,n}
+            \vdots \\
+            S_{m,n}
         \end{bmatrix} = 
@@
             C_{2,1} & C_{2,2} &\dots & C_{2,n-1}\\
-            C_{3,1} & C_{3,2} &\dots &C_{3,n-1}\\
-        \end{bmatrix} + 
+            \vdots & \vdots  & \ddots &\vdots\\
+            C_{m,1} & C_{m,2} &\dots &C_{m,n-1}\\
+        \end{bmatrix} \cdot
         \begin{bmatrix}
@@
             S_{2,1} & S_{2,2} & \dots & S_{2,n-1}\\
-            S_{3,1} & S_{3,2} & \dots & S_{3,n-1}\\
+            \vdots & \vdots  & \ddots &\vdots\\
+            S_{m,1} & S_{m,2} & \dots & S_{m,n-1}\\
         \end{bmatrix}+
@@
             \epsilon_{2}\\
-            \epsilon_{3}
+            \vdots\\
+            \epsilon_{n}\\
         \end{bmatrix}$%
@@
 \end{equation*}
-Each time series $S_n$ $\forall i \in \{1,2,3\}$ describes the n-th data point for the input time series i. The term $\epsilon$ is added to approximate an error, which is 
-induced by the model.\\
-We need to assure that our model is stationary, which means that the the stochastic process doesn't change over time. \textcolor{red}{WHY?}.
-Our data is stationary if we are able to assure this for all our time series data inputs:
+The $\epsilon$ vector approximates an error, which is induced by the model. One main benefit of the VAR model is that it also assumes that $S_{i,j}$ is also influenced by $S_{k,l}$ 
+$i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$.
+In our case, we have three input time series, the students, salary and graduates data from the Winter Semester 2007 to the Summer Semester 2021. This model has one hyperparameter which describes
+the amount of past data that should be included in our models prediction. 
+The vector autoregressive model heavily depends on the stationarity assumption which ensures that the stochastical process stays the same over time. 
+In this context, this is assured by the following three axioms
 \begin{enumerate}[label=\Roman*]
@@
 \end{enumerate}
-
-After we ensured, that our data is stationary, we used a PACF plot to find lags which could be statistically significant. 
-The main reason why we choose PACF in contrast to ACF is that we only want to look at direct influences from $S_{n-1}$ to $S_{n}$ and not from $S_{n-3}$ to $S_n$.
-\begin{figure}
-    \includegraphics{fig/pacf_plot.pdf}
-    \caption{PACF Plot for our VAR Model}
-\end{figure}
-\textcolor{red}{// CONCLUSION FOR PACF // }\\
-
-After we found the optimal lag configuration of our model, we used a Granger Causality Test, to statistically prove that the salary time series really influences the student decision.
-The main idea of Granger Causality is that we are able to predict the data point $S_{i,n}$ directly with one or multiple other time serieses. The Granger Causality Test uses exactly this
-idea, by inserting data points $S*:=S_{j,m}$ for $j\in i:j\neq i$  and $m<n$. After we inserted a couple of terms we do a t-Test and an F-Test on this new $S*$ and as soon as both tests
-return statistical significance for the same S* we can assure that the three time series are Granger Causal.
+We assume that our time series is mostly stationary, and this is enough for our case because we are only predicting 2 data points in the future.
+In order to find the optimal hyperparameter we used two well-established methods, the PACF and also an information-based criteria particularly the AIC and 
+BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ onto $S_{n}$, and then it builds a $(1-\alpha)$ confidence interval. We choose $\alpha = 5\%$ and get
+get a 95\% Confidence Interval 
+\begin{figure*}
+    \includegraphics{fig/prediction_lag_1.pdf}
+    \caption{Prediction based on lag order 1}
+\end{figure*}
+As you can see, the prediction of the average salary after tax is almost linear, but the true average salary after tax has a totally different shape.
+After further analysis, we decided to use the hyperparameter p based
+on an information criterion namely AIC and BIC.
+The abbreviation AIC stands for Akaike information criterion and tries to find an optimum between the complexity of the model and fitting the data, which should pretend overfitting.
+BIC is an abbreviation for Bayesian Information Criterion and the main difference between AIC and BIC is that BIC penalizes more complex models, which overall leads to less complicated models than the ones from AIC.
+The optimal lag based on AIC is 5, while for BIC it is 1, which aligns with the PACF.
+\begin{figure*}
+    \includegraphics{fig/prediction_lag_5.pdf}
+    \caption{Prediction based on lag order 5}
+\end{figure*}
+As you can see the lag order 5 mimics the average salary after tax better than the lag order 1.\\
+Another quite important aspect in the correlation analysis is the Granger Causality Test which we did on the output. 
+Granger Causal implies that one time series is the result of another shifted time series.
+In our case, we try to find out if e.g. the salary time series is just a shifted version of the students or the graduates time series.
+The Granger causality test does exactly this, it uses the time series $S_i$ and tries to insert data points from $S_j$ or $S_k$ $i \neq j\neq k$
+to predict the outcome of the model. After inserting the data points the test performs a t-test and a F-test, if both of these tests return statistically significant,
+then we showed that $S_i$ could be expressed by a shifted input time series $S_j$ with $j\neq i$
 

### 2024-01-28 src/model.py
refactor model creation

@@ -1,0 +1,57 @@
+"""This module contains all methods to build the model and to make predictions."""
+
+import numpy as np
+import pandas as pd
+from statsmodels.tsa.api import VAR
+from statsmodels.tsa.vector_ar.var_model import VARResultsWrapper
+from src.constants import YEARS
+
+
+def fitVarModel(students: np.ndarray, salary: np.ndarray, graduates: np.ndarray, lags: int) -> VARResultsWrapper:
+    """
+    Fits a Vector Autoregression (VAR) model to the given data.
+
+    Parameters:
+    students (array-like): The number of students for each year.
+    salary (array-like): The average salary for each year.
+    graduates (array-like): The number of graduates for each year.
+    lags (int): The maximum number of lags to use in the model.
+
+    Returns:
+    VARResultsWrapper: The results of the fitted VAR model.
+    """
+    dataset = np.column_stack((students, salary, graduates.repeat(2)))
+    dataset = pd.DataFrame(dataset, index=YEARS[3:-4], columns=['students', 'salary', 'graduates'])
+    model = VAR(dataset)
+    results = model.fit(maxlags=lags)
+    return results
+
+
+def predict(results: VARResultsWrapper, steps: int) -> np.ndarray:
+    """
+    Makes a prediction using the fitted VAR model.
+
+    Parameters:
+    results (VARResultsWrapper): The results of the fitted VAR model.
+    steps (int): The number of steps ahead to forecast.
+
+    Returns:
+    array: The forecasted values.
+    """
+    prediction, _, _ = predictWithStderr(results, steps)
+    return prediction
+
+
+def predictWithStderr(results: VARResultsWrapper, steps: int) -> tuple:
+    """
+    Makes a prediction using the fitted VAR model and also returns the standard error.
+
+    Parameters:
+    results (VARResultsWrapper): The results of the fitted VAR model.
+    steps (int): The number of steps ahead to forecast.
+
+    Returns:
+    tuple: A tuple containing the forecasted values, the lower bound of the forecast interval, and the upper bound of the forecast interval.
+    """
+    forecast = results.forecast_interval(results.endog, steps=steps, alpha=0.05)
+    return forecast

### 2024-01-28 src/model.py
refactor exp_02

@@ -27,22 +27,23 @@ def fitVarModel(students: np.ndarray, salary: np.ndarray, graduates: np.ndarray,
     return results
 
 
-def predict(results: VARResultsWrapper, steps: int) -> np.ndarray:
+def fitVarModelCompleteDataset(dataset: np.ndarray, lags: int) -> VARResultsWrapper:
     """
-    Makes a prediction using the fitted VAR model.
+    Fits a Vector Autoregression (VAR) model to the given data.
 
     Parameters:
-    results (VARResultsWrapper): The results of the fitted VAR model.
-    steps (int): The number of steps ahead to forecast.
+    dataset (array-like): The dataset to fit the model to.
+    lags (int): The maximum number of lags to use in the model.
 
     Returns:
-    array: The forecasted values.
+    VARResultsWrapper: The results of the fitted VAR model.
     """
-    prediction, _, _ = predictWithStderr(results, steps)
-    return prediction
+    model = VAR(dataset)
+    results = model.fit(maxlags=lags)
+    return results
 
 
-def predictWithStderr(results: VARResultsWrapper, steps: int) -> tuple:
+def predict(results: VARResultsWrapper, steps: int) -> tuple[np.ndarray, np.ndarray, np.ndarray]:
     """
     Makes a prediction using the fitted VAR model and also returns the standard error.
 
@@ -54,4 +55,19 @@ def predictWithStderr(results: VARResultsWrapper, steps: int) -> tuple:
     tuple: A tuple containing the forecasted values, the lower bound of the forecast interval, and the upper bound of the forecast interval.
     """
     forecast = results.forecast_interval(results.endog, steps=steps, alpha=0.05)
+    return forecast
+
+
+def predictWithData(results: VARResultsWrapper, data, steps: int) -> tuple[np.ndarray, np.ndarray, np.ndarray]:
+    """
+    Makes a prediction using the fitted VAR model and also returns the standard error.
+
+    Parameters:
+    results (VARResultsWrapper): The results of the fitted VAR model.
+    steps (int): The number of steps ahead to forecast.
+
+    Returns:
+    tuple: A tuple containing the forecasted values, the lower bound of the forecast interval, and the upper bound of the forecast interval.
+    """
+    forecast = results.forecast_interval(data, steps=steps, alpha=0.05)
     return forecast

### 2024-01-28 src/dataLoading.py
refactor brutto

@@ -3,7 +3,6 @@
 import pandas as pd
 import numpy as np
 from .constants import GRADUATES_PATH, STUDENTS_PATH, SALARY_PATH, INFLATION_PATH, SEMESTERS
-import os 
 
 
 def getStudents() -> pd.DataFrame:
@@ -98,15 +97,15 @@ def getTotalStudentsFor(courses: list, years: list[str]) -> np.array:
     return studentsInCourseHF.sum(axis=0) + studentsInCourseNF.sum(axis=0)
 
 
-def getBruttoSalary(sector) -> np.array:
-    salaries = getSalaries()    
+def getGrossSalary(sector) -> np.array:
+    salaries = getSalaries()
     SALARY_YEARS = salaries.index.levels[1]
     QUARTALS = ['1. Quartal', '2. Quartal', '3. Quartal', '4. Quartal']
 
-    bruttoSalaryNP = salaries.loc[(sector, SALARY_YEARS, QUARTALS), ('Insgesamt', 'Insgesamt', 'Durchschnittliche Bruttomonatsverdienste', 'EUR')].to_numpy(dtype=int) # noqa: E501
-    bruttoSalaryMeanForYears = bruttoSalaryNP.reshape(-1, 2).mean(axis=1)
-    bruttoSalary = bruttoSalaryMeanForYears.reshape(-1,30).mean(axis=0)
-    return bruttoSalary
+    grossSalaryNP = salaries.loc[(sector, SALARY_YEARS, QUARTALS), ('Insgesamt', 'Insgesamt', 'Durchschnittliche Bruttomonatsverdienste', 'EUR')].to_numpy(dtype=int)  # noqa: E501
+    grossSalaryMeanForYears = grossSalaryNP.reshape(-1, 2).mean(axis=1)
+    grossSalary = grossSalaryMeanForYears.reshape(-1, 30).mean(axis=0)
+    return grossSalary
 
 
 def getGraduatesInBwFor(years: list) -> np.array:
@@ -120,11 +119,11 @@ def getAllGraduatesYears() -> list[str]:
     return graduates.index.tolist()
 
 
-def getInflationAdjustedBruttoSalary(sector) -> np.array:
+def getInflationAdjustedGrossSalary(sector) -> np.array:
     inflation = getInflation()
     inflation = inflation.loc[:, ('Veränderungsrate zum Vorjahr', 'Prozent')].to_numpy(dtype=float)
 
-    bruttoSalary = getBruttoSalary(sector)
[…]

### 2024-01-28 src/dataLoading.py
reduce prints

@@ -4,6 +4,8 @@ import pandas as pd
 import numpy as np
 from .constants import GRADUATES_PATH, STUDENTS_PATH, SALARY_PATH, INFLATION_PATH, SEMESTERS
 
+DEBUG = False
+
 
 def getStudents() -> pd.DataFrame:
     """Load student data from excel file into a pandas DataFrame."""
@@ -142,9 +144,10 @@ def getInflationAdjustedGrossSalaries(sectors: list) -> np.array:
     inflation = inflation.loc[:, ('Veränderungsrate zum Vorjahr', 'Prozent')].to_numpy(dtype=float)
 
     cummulativeInflation = np.cumprod(1 + inflation / 100)
-    print('Create cummulative inflation:')
-    for a, b in zip(inflation, cummulativeInflation):
-        print(f'{a} -> {b}')
+    if DEBUG:
+        print('Create cummulative inflation:')
+        for a, b in zip(inflation, cummulativeInflation):
+            print(f'{a} -> {b}')
 
     # Initialize an array to store the sum of salaries for all sectors
     sum_grossInflationSalary = np.zeros_like(cummulativeInflation).repeat(2)
@@ -154,9 +157,10 @@ def getInflationAdjustedGrossSalaries(sectors: list) -> np.array:
         grossSalary = getGrossSalary(sector)
         grossInflationSalary = grossSalary / cummulativeInflation.repeat(2)
         sum_grossInflationSalary += grossInflationSalary  # Sum the salaries
-        print('\nAdjust salary for inflation:')
-        for a, b in zip(grossSalary, grossInflationSalary):
-            print(f'{a} -> {b}')
+        if DEBUG:
+            print('\nAdjust salary for inflation:')
+            for a, b in zip(grossSalary, grossInflationSalary):
+                print(f'{a} -> {b}')
 
     # Calculate the average inflation-adjusted salary
     avg_grossInflationSalary = sum_grossInflationSalary / len(sectors)

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
abstract and introduction

--- 2024-01-25 doc/DataLiteracyStudentProject/paper.tex
@@
 \twocolumn[
-\icmltitle{My Data Literacy Project\\ (Replace this with your Project Title)}
+\icmltitle{All just Greedy for Money? - An Prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field}
 
@@
 \icmlaffiliation{first}{Matrikelnummer 5977981, abdallah.abdul-latif@student.uni-tuebingen.de, MSc Computer Science}
-\icmlaffiliation{second}{Matrikelnummer 12345678, first.last@student.uni-tuebingen.de, MSc Computer Science}
+\icmlaffiliation{second}{Matrikelnummer 4189024, lisa-maria.fritsch@student.uni-tuebingen.de, MSc Computer Science}
 \icmlaffiliation{third}{Matrikelnummer 12345678, first.last@student.uni-tuebingen.de, MSc Media Informatics}
@@
 \begin{abstract}
-As often mentioned in the news and in the daily life, people suggest that a higher expected salary is a strong pull factor for students 
-choosing their subject. We were curious if this is particullarily true for the scientific based subjects, because a majority of them have a 
-pretty high expected salary.
-We used the students data from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics University of Tübingen},
-the high-school graduates data from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg}
-and also the salary data from 2005 to 2023 from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt}.
-We used a vector autoregressive model, to build a time series forecast and analysed the outcomes to find an optimal configuration for this type of model in our context.
-\textcolor{red}{result.}
+This paper examines whether the expected salary plays a role in which subject graduates choose to study.
+We have limited ourselves to the STEM (science, technology, engineering and maths) field.
+We analyse the influence of the three parameters: number of Graduates, of Students and Expected Salary, have on each other.
+Therefore we use the Vector autoregressive model and a PACF.
+This showed that the expected salary has an \textcolor{red}{or has no?} influence on how many school leavers decide to study in the STEM field.
+We were also able to make the prediction that the expected salary and the number of students in the next two years will be similar to the current situation, but slightly reduced.
+If the salary is increased or decreased in the analysis, it can be seen that the number of students goes up or down accordingly.(\textcolor{red}{or else?})
 %Motivate the problem, situation or topic you decided to work on. 
@@
 \section{Introduction}\label{sec:intro}
-In the public eye, there is a strong correlation between the expected salary and the subject enrollment of students. We are curious if this statement is true or just a biased
-perception of the society. Our main focus is on the University of Tübingen and mainly on the STEM (scientific engineering technical and mathematical) subjects, because those degrees 
-are likely to have a pretty high expected salary after graduating. Furthermore we needed data of high-school graduates in Baden-Würrtemberg, to find a possible corellation between
-the amount of graduates and the new enrollments in the stem subjects. We just focused on Baden-Württemberg, this introduces a bias to our prediction, but we assumed that the 
-majority of students which are going to study in Tübingen come from Baden-Württemberg. The third datatset we used is about the salaries from 2005 to 2023 in the main 
-industrial sectors in Germany. The main focus of this paper is on a analysis of a vetor autoregressive model, to predict the upcoming numbers of students in the next semesters. We
-used various techniques to clean our data like removing the inflation or making our time series stationary to ensure that the predictions don't detect patterns which are inferred by 
-some noise in the data. Nevertheless we also used basic data science techniques to get a better understanding of the dirstributions and trends over the past years.
+The trend in recent years shows that more and more pupils are taking their high-school diploma and then going to university.
+Many assume that one of the reasons for this is better salary prospects.
+We are interested in whether this is actually true.
+To investigate this, we used three data sets.
+The first data set is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}.
+This can be used to determine how many students there are in which semester and in which subject in Tübingen.
+We limit ourselves to the STEM field because this area promises a high and growing salary.
+We also use the data from Tübingen because the STEM field is well represented here.
+This university can therefore be used as a representative.
+However, this can lead to a bias in our prediction.
+We obtained our second data set from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg}.
+It contains the number of high-school graduates from Baden-Württemberg for the past years.
+We restrict ourselves to Baden-Württemberg, as it can be assumed that most students in Tübingen come from Baden-Württemberg.
+As this does not apply to all students, there is also a bias.
+The last data set we need was obtained from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt}.
+It contains the salaries of different industrial sectors in Germany.
+Here we have used the sectors that are relevant for the STEM field.
+We extracted the relevant data from the three datasets and analysed them with the vector autoregressive model and a PACF (\Cref{sec:methods}).
+This results in a model that can be used to predict how many students there will be in the STEM field in the next two years and how the expected salary will develop.
+We can also analyse the interplay of the three parameters and predict what will happen if we modify the salary.
+(\Cref{sec:results})
 

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
revised abstract

@@
 
-% if you use cleveref..
+% if you use cleveref.
 \usepackage[capitalize,noabbrev]{cleveref}
@@
 \twocolumn[
-\icmltitle{All just Greedy for Money? - An Prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field}
+\icmltitle{All just Greedy for Money? - A prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field}
 
@@
 
-%Put your abstract here. Abstracts typically start with a sentence motivating why the subject is interesting. Then mention the data, methodology or methods you are working with, and describe results. 
+%Put your abstract here.
+Abstracts typically start with a sentence motivating why the subject is interesting.
+Then mention the data, methodology, or methods you are working with, and describe results.
 \begin{abstract}
-This paper examines whether the expected salary plays a role in which subject graduates choose to study.
-We have limited ourselves to the STEM (science, technology, engineering and maths) field.
-We analyse the influence of the three parameters: number of Graduates, of Students and Expected Salary, have on each other.
-Therefore we use the Vector autoregressive model and a PACF.
-This showed that the expected salary has an \textcolor{red}{or has no?} influence on how many school leavers decide to study in the STEM field.
-We were also able to make the prediction that the expected salary and the number of students in the next two years will be similar to the current situation, but slightly reduced.
-If the salary is increased or decreased in the analysis, it can be seen that the number of students goes up or down accordingly.(\textcolor{red}{or else?})
-%Motivate the problem, situation or topic you decided to work on. 
-% Describe why it matters (is it of societal, economic, scientific value?). 
+This paper investigates the role of expected salary in influencing the choice of academic majors among
+students in the STEM field at the University of Tübingen.
+The study focuses on the interplay among three key parameters: the enrollment of students, the expected salary, and the number of graduates.
+Employing a Vector Autoregressive model, we can confidently predict the number of students for the upcoming semesters
+and show that the expected salary plays a key role in making that prediction.
+Furthermore, by manipulating the salary variable in our analysis, we observe a corresponding increase or decrease in the number of students.
+This implies a sensitivity in enrollment trends to changes in expected salary, highlighting the potential impact of financial considerations on academic choices.
+
+% Motivate the problem, situation, or topic you decided to work on. 
+% Describe why it matters (is it of societal, economic, or scientific value?). 
 % Outline the rest of the paper (use references, e.g.~to \Cref{sec:methods}: 
-% What kind of data you are working with, how you analyse it, and what kind of conclusion you reached. 
+% What kind of data you are working with, how you analyze it, and what kind of conclusion you reached. 
 % The point of the introduction is to make the reader want to read the rest of the paper.

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
update result

--- 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
@@
 % You can use the \verb|\citep| and \verb|\citet| commands for this purpose \citep{mackay2003information}.
+
 \section{Data and Methods}\label{sec:methods}
@@
 get a 95\% Confidence Interval 
-\begin{figure*}
-    \includegraphics{fig/prediction_lag_1.pdf}
-    \caption{Prediction based on lag order 1}
-\end{figure*}
-As you can see, the prediction of the average salary after tax is almost linear, but the true average salary after tax has a totally different shape.
-After further analysis, we decided to use the hyperparameter p based
-on an information criterion namely AIC and BIC.
-The abbreviation AIC stands for Akaike information criterion and tries to find an optimum between the complexity of the model and fitting the data, which should pretend overfitting.
+\begin{table}[H]
+    \centering
+    \begin{tabular}{c|c|c|c}
+        & students & salary & graduates \\
+        \hline
+        Prediciton 1 & 1.25723662 & 1.27163776 & 3.94417629 \\
+        \hline
+        Prediciton 2 & 4.84256257 & 0.11922267 & 1.72230067
+    \end{tabular}
+    \caption{Percentage difference between true value and predicted value for students, salary and graduates for lag 1}
+\end{table}
+As you can see, the percent difference between the true values especially for the students are quite high, becasue of that we tried the AIC and BIC to find a better lag for the 
+hyperparameter p.
+The abbreviation AIC stands for Akaike Information Criterion and tries to find an optimum between the complexity of the model and fitting the data, which should pretend overfitting.
 BIC is an abbreviation for Bayesian Information Criterion and the main difference between AIC and BIC is that BIC penalizes more complex models, which overall leads to less complicated models than the ones from AIC.
-The optimal lag based on AIC is 5, while for BIC it is 1, which aligns with the PACF.
-\begin{figure*}
-    \includegraphics{fig/prediction_lag_5.pdf}
-    \caption{Prediction based on lag order 5}
-\end{figure*}
-As you can see the lag order 5 mimics the average salary after tax better than the lag order 1.\\
+The optimal lag based on AIC is 6, while for BIC it is 1, which aligns with the PACF.
+\begin{table}[H]
+    \centering
+    \begin{tabular}{c|c|c|c}
+        & students & salary & graduates \\
+        \hline
+        Prediciton 1 & 0.48413863 & 0.41885829 & 0.11117345 \\
+        \hline
+        Prediciton 2 & 1.34596809 & 0.30583828 & 0.93859448
+    \end{tabular}
+    \caption{Percentage difference between true value and predicted value for students, salary and graduates for lag 5}
+\end{table}
+As you can see there is a major improvement in the prediction for the students in comparison to lag order 1. We decided to use lag 5, because the main focus of our 
+of our research is to predict the students based on salary and graduates.
 Another quite important aspect in the correlation analysis is the Granger Causality Test which we did on the output. 
@@
 \section{Results}\label{sec:results}
+The hyperparameter optimization is always a trade-off between several optimal values for different parameters. 
+Our main goal was to predict the students who enroll in study subjects in the STEM sector when using the salary and the
+graduates of the given years as prior knowledge. For this reason, we used a lag of 5, because the overall percentage error
+was the lowest for the students.
+\begin{figure}[ht]
+    \includegraphics{fig/prediction_5.pdf}
+    \caption{VAR Model Prediction with hyperparameter p=5}
+\end{figure}
+During our analysis, we found out, that the salary has a major influence on our prediction. If our model works correctly, then the amount
+of predicted students should go down, because fewer students imply that there is a lack of qualified workers on the job market.
+This on the other hand leads to an increase in the salary. Now we manipulated the salary data with a fixed $\alpha\in \mathbb{N}$ to increase
+the whole time series by a constant factor. We used this approach because the vector autoregressive model internally uses Mean Squared Error as loss function.
+The Mean Squared Error squares the difference between the data points, which leads to a stronger influence of large differences.
+This high error leads to a wrong minimization, which gives wrong coefficients.
+Another more important point is that we don't want to violate stationarity, which ensures that our model predicts in a good manner.\\
+\begin{table}[H]
+    \centering
+    \begin{tabular}{c|c|c|c}
+        & students & salary & graduates\\
+        \hline
+        students & 1.000000 & -0.115442 &  0.111069\\
+        \hline
+        salary & -0.115442 & 1.000000 & -0.902321\\
+        \hline
+        graduates & 0.111069 & -0.902321 &  1.000000\\
+    \end{tabular}
+    \caption{Correlation Matrix of Var(5) model}
+\end{table}
 
-In this section outline your results. At this point, you are just stating the outcome of your analysis. 
-You can highlight important aspects (``we observe a significantly higher value of $x$ over $y$''), but leave interpretation and opinion to the next section.
-This section absoultely \emph{has} to include at least two figures.
+As you can see the correlation matrix, tells us that the model also takes the relationship into account, that a 
+manipulated salary has an inverse relationship to the amount of students.
 
-\begin{figure*}
-    \includegraphics{fig/params.pdf}
-    \caption{This is a figure caption. Write this, if we end up using that fig ;-)}
-\end{figure*}
+\begin{figure}[H]
+    \centering
+    \includegraphics{fig/prediction_modified_10\%_increase_lags_5.pdf}
+    \caption{Increased salary by 10 \% in Var(5) model}
+\end{figure}
 
+% Use this section to briefly summarize the entire text. Highlight limitations and problems, 
+% but also make clear statements where they are possible and supported by the analysis. 
 \section{Discussion \& Conclusion}\label{sec:conclusion}
-
-Use this section to briefly summarize the entire text. Highlight limitations and problems, but also make clear statements where they are possible and supported by the analysis.
 

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
update introduction

--- 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
@@
 
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+%\printAffiliationsAndNotice{}  % leave blank if no need to mention the equal contribution
 \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
@@
 Abstracts typically start with a sentence motivating why the subject is interesting.
-Then mention the data, methodology, or methods you are working with, and describe results.
+Then mention the data, methodology, or methods you are working with, and describe the results.
 \begin{abstract}
@@
 \section{Introduction}\label{sec:intro}
-The trend in recent years shows that more and more pupils are taking their high-school diploma and then going to university.
-Many assume that one of the reasons for this is better salary prospects.
-We are interested in whether this is actually true.
-To investigate this, we used three data sets.
-The first data set is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}.
-This can be used to determine how many students there are in which semester and in which subject in Tübingen.
-We limit ourselves to the STEM field because this area promises a high and growing salary.
-We also use the data from Tübingen because the STEM field is well represented here.
-This university can therefore be used as a representative.
-However, this can lead to a bias in our prediction.
-We obtained our second data set from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg}.
-It contains the number of high-school graduates from Baden-Württemberg for the past years.
-We restrict ourselves to Baden-Württemberg, as it can be assumed that most students in Tübingen come from Baden-Württemberg.
-As this does not apply to all students, there is also a bias.
-The last data set we need was obtained from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt}.
-It contains the salaries of different industrial sectors in Germany.
-Here we have used the sectors that are relevant for the STEM field.
-We extracted the relevant data from the three datasets and analysed them with the vector autoregressive model and a PACF (\Cref{sec:methods}).
-This results in a model that can be used to predict how many students there will be in the STEM field in the next two years and how the expected salary will develop.
-We can also analyse the interplay of the three parameters and predict what will happen if we modify the salary.
-(\Cref{sec:results})
+In recent years, a growing number of high school graduates have chosen the path of higher education.
+Science, technology, engineering, and mathematics, known as STEM, make up a large proportion of all students, although
+this academic career is known for being one of the most demanding.
+This paper aims to explore the potential influence of financial incentives.
+
+% This is some text that I like but does not fit here. Maybe I find some place to put it.
+
+% Our central contribution is a Vector autoregressive model (VAR) (\Cref{sec:methods}) that, combined with a
+% Partial autocorrelation function (PACF) (\Cref{sec:methods}), predicts the salary and number
+% of students enrolled in STEM courses at the University of Tübingen.
+
+The analysis relies on three underlying datasets.
+The first dataset captures the historical enrollment figures of students at the University of Tübingen. 
+It contains the total number of students for each semester, for all study programs. This dataset is maintained and published by the
+\href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}.
+The University of Tübingen makes a good candidate to discover potential structure in the number
+of enrollments because STEM programs are strongly represented here. % TODO: I think mentioning the potential risk of BIAS could be in the Discussion-part
+To quantify the salary expectations, a dataset from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt} is used.
+It contains the salary trend for all major industrial sectors in Germany. This dataset does not adjust for inflation,
+therefore an additional \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=61121-0001&bypass=true&levelindex=0&levelid=1706446467491#abreadcrumb}{dataset} from the Deutsches Statistisches Bundesamt is used to account for that.
+It contains the inflation rate in Germany for each year.
+The number of students is also limited by a further parameter. Namely the quantity of people eligible for an academic career.
+The \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} provides
+data about high school graduates in Baden-Württemberg for every year. We will not take other states into account, under
+the assumption that the number of students leaving Baden-Württemberg, to study elsewhere, cancels out with others moving to the University of Tübingen.
+We extracted the relevant parts of each dataset, to suit our analysis with a VAR model (\Cref{sec:methods}).
 
 % In this section, describe \emph{what you did}. Roughly speaking, 
-% explain what data you worked with, how or from where it was collected, it's structure and size.
-%  Explain your analysis, and any specific choices you made in it. Depending on the nature of your project,
-% you may focus more or less on certain aspects. If you collected data yourself, explain the collection process in detail.
+% explain what data you worked with, how or from where it was collected, its structure, and size.
+% Explain your analysis and any specific choices you made in it. Depending on the nature of your project,
+% you may focus more or less on certain aspects. If you collect data yourself, explain the collection process in detail.
 % If you downloaded data from the net, show an exploratory analysis that builds
@@
 \section{Data and Methods}\label{sec:methods}
+
 In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of theUniversity of Tübingen}., which contains the

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
current state of the paper
added a comment

--- 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
@@
 \twocolumn[
-\icmltitle{My Data Literacy Project\\ (Replace this with your Project Title)}
+\icmltitle{All just Greedy for Money? - A prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field}
 
@@
 
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
+%\printAffiliationsAndNotice{}  % leave blank if no need to mention the equal contribution
 \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
 
-%Put your abstract here. Abstracts typically start with a sentence motivating why the subject is interesting. Then mention the data, methodology or methods you are working with, and describe results. 
+%Put your abstract here.
+Abstracts typically start with a sentence motivating why the subject is interesting.
+Then mention the data, methodology, or methods you are working with, and describe the results.
 \begin{abstract}
-As often mentioned in the news and in the daily life, people suggest that a higher expected salary is a strong pull factor for students 
-choosing their subject. We were curious if this is particullarily true for the scientific based subjects, because a majority of them have a 
-pretty high expected salary.
-We used the students data from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics University of Tübingen},
-the high-school graduates data from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg}
-and also the salary data from 2005 to 2023 from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt}.
-We used a vector autoregressive model, to build a time series forecast and analysed the outcomes to find an optimal configuration for this type of model in our context.
-\textcolor{red}{result.}
-%Motivate the problem, situation or topic you decided to work on. 
-% Describe why it matters (is it of societal, economic, scientific value?). 
+This paper investigates the role of expected salary in influencing the choice of academic majors among
+students in the STEM field at the University of Tübingen.
+The study focuses on the interplay among three key parameters: the enrollment of students, the expected salary, and the number of graduates.
+Employing a Vector Autoregressive model, we can confidently predict the number of students for the upcoming semesters
+and show that the expected salary plays a key role in making that prediction.
+Furthermore, by manipulating the salary variable in our analysis, we observe a corresponding increase or decrease in the number of students.
+This implies a sensitivity in enrollment trends to changes in expected salary, highlighting the potential impact of financial considerations on academic choices.
+
+% Motivate the problem, situation, or topic you decided to work on. 
+% Describe why it matters (is it of societal, economic, or scientific value?). 
 % Outline the rest of the paper (use references, e.g.~to \Cref{sec:methods}: 
-% What kind of data you are working with, how you analyse it, and what kind of conclusion you reached. 
+% What kind of data you are working with, how you analyze it, and what kind of conclusion you reached. 
 % The point of the introduction is to make the reader want to read the rest of the paper.
@@
 \section{Introduction}\label{sec:intro}
-In the public eye, there is a strong correlation between the expected salary and the subject enrollment of students. We are curious if this statement is true or just a biased
-perception of the society. Our main focus is on the University of Tübingen and mainly on the STEM (scientific engineering technical and mathematical) subjects, because those degrees 
-are likely to have a pretty high expected salary after graduating. Furthermore we needed data of high-school graduates in Baden-Würrtemberg, to find a possible corellation between
-the amount of graduates and the new enrollments in the stem subjects. We just focused on Baden-Württemberg, this introduces a bias to our prediction, but we assumed that the 
-majority of students which are going to study in Tübingen come from Baden-Württemberg. The third datatset we used is about the salaries from 2005 to 2023 in the main 
-industrial sectors in Germany. The main focus of this paper is on a analysis of a vetor autoregressive model, to predict the upcoming numbers of students in the next semesters. We
-used various techniques to clean our data like removing the inflation or making our time series stationary to ensure that the predictions don't detect patterns which are inferred by 
-some noise in the data. Nevertheless we also used basic data science techniques to get a better understanding of the dirstributions and trends over the past years.
+In recent years, a growing number of high school graduates have chosen the path of higher education.
+Science, technology, engineering, and mathematics, known as STEM, make up a large proportion of all students, although
+this academic career is known for being one of the most demanding.
+This paper aims to explore the potential influence of financial incentives.
+
+% This is some text that I like but does not fit here. Maybe I find some place to put it.
+
+% Our central contribution is a Vector autoregressive model (VAR) (\Cref{sec:methods}) that, combined with a
+% Partial autocorrelation function (PACF) (\Cref{sec:methods}), predicts the salary and number
+% of students enrolled in STEM courses at the University of Tübingen.
+
+The analysis relies on three underlying datasets.
+The first dataset captures the historical enrollment figures of students at the University of Tübingen. 
+It contains the total number of students for each semester, for all study programs. This dataset is maintained and published by the
+\href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}.
+The University of Tübingen makes a good candidate to discover potential structure in the number
+of enrollments because STEM programs are strongly represented here. % TODO: I think mentioning the potential risk of BIAS could be in the Discussion-part
+To quantify the salary expectations, a dataset from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt} is used.
+It contains the salary trend for all major industrial sectors in Germany. This dataset does not adjust for inflation,
+therefore an additional \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=61121-0001&bypass=true&levelindex=0&levelid=1706446467491#abreadcrumb}{dataset} from the Deutsches Statistisches Bundesamt is used to account for that.
+It contains the inflation rate in Germany for each year.
+The number of students is also limited by a further parameter. Namely the quantity of people eligible for an academic career.
+The \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} provides
+data about high school graduates in Baden-Württemberg for every year. We will not take other states into account, under
+the assumption that the number of students leaving Baden-Württemberg, to study elsewhere, cancels out with others moving to the University of Tübingen.
+We extracted the relevant parts of each dataset, to suit our analysis with a VAR model (\Cref{sec:methods}).
 
 % In this section, describe \emph{what you did}. Roughly speaking, 
-% explain what data you worked with, how or from where it was collected, it's structure and size.
-%  Explain your analysis, and any specific choices you made in it. Depending on the nature of your project,
-% you may focus more or less on certain aspects. If you collected data yourself, explain the collection process in detail.
+% explain what data you worked with, how or from where it was collected, its structure, and size.
+% Explain your analysis and any specific choices you made in it. Depending on the nature of your project,
+% you may focus more or less on certain aspects. If you collect data yourself, explain the collection process in detail.
 % If you downloaded data from the net, show an exploratory analysis that builds
@@
 \section{Data and Methods}\label{sec:methods}
-In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}., which contains the
+% Mabe focus here on what we did with the data
+% I.e. inflation adjustment
+% flattening G8 G9 conflict
+
+In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of theUniversity of Tübingen}., which contains the
+
 total amount of students from the winter semester 2005/2006 to the winter semester 2023/2024. The data was collected from the enrollments of students, which registered 
@@
         \hline
-        Prediciton 1 & 0.48413863 & 0.41885829 & 0.11117345 \\
+        Prediciton 1 & 0.48413863\% & 0.41885829\% & 0.11117345\% \\
         \hline
-        Prediciton 2 & 1.34596809 & 0.30583828 & 0.93859448
+        Prediciton 2 & 1.34596809\% & 0.30583828\% & 0.93859448\%
     \end{tabular}
@@
 was the lowest for the students.
-\begin{figure}[ht]
+\begin{figure*}
     \includegraphics{fig/prediction_5.pdf}
     \caption{VAR Model Prediction with hyperparameter p=5}
-\end{figure}
+\end{figure*}
 During our analysis, we found out, that the salary has a major influence on our prediction. If our model works correctly, then the amount

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
Bug fix
Introductino

@@
 
-% This is some text that I like but does not fit here. Maybe I find some place to put it.
-
-% Our central contribution is a Vector autoregressive model (VAR) (\Cref{sec:methods}) that, combined with a
-% Partial autocorrelation function (PACF) (\Cref{sec:methods}), predicts the salary and number
-% of students enrolled in STEM courses at the University of Tübingen.
-
 The analysis relies on three underlying datasets.
@@
 the assumption that the number of students leaving Baden-Württemberg, to study elsewhere, cancels out with others moving to the University of Tübingen.
-We extracted the relevant parts of each dataset, to suit our analysis with a VAR model (\Cref{sec:methods}).
+
+Our central contribution is a Vector autoregressive model (VAR) discussed in \Cref{sec:methods} that is built using just mentioned data. Combined with a
+Partial autocorrelation function (PACF) the optimal value for lag
+is determined, see \Cref{sec:methods} \textcolor{red}{subsection}, to ensure that relevant structure in the data is captured by the model.
+After verifying that the VAR model functions as expected and testing the model on a test set, \Cref{sec:methods} \textcolor{red}{subsection}, we
+can predict all three input time series. As we are only interested in the number of students enrolled and the development of the salary, our research is 
+focused on these two instances.
+This allows us to conclude that salary has a significant impact on the amount of STEM students. Further, we can model different salary development scenarios, 
+and see how the students would likely react (\Cref{sec:results}).
 
@@
 
-In our research we used three datasets, which are divided as follows: The first dataset is from the \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of theUniversity of Tübingen}., which contains the
-
-total amount of students from the winter semester 2005/2006 to the winter semester 2023/2024. The data was collected from the enrollments of students, which registered 
-for a study in Tübingen.
-The second dataset is from the \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} which contains data about the high-school graduates from 1970 to 2023.
-Our third and last dataset is from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt} and contains the salaries of the large industrial sectors in Germany from 2005 to 2021.
-One main problem with the mapping between university subjects and actual jobs is that there exists no perfect one-to-one mapping. 
-We also removed the inflation from the salaries to get the real wage, because the inflation could 
-cause a false pattern detection in our autoregressive model. 
-To reduce the amount of bias introduced in this step, we also used the Inflation data from the Deutsches Statistisches Bundesamt over the years.\\
+Each data set used in the analysis had to be processed, to work for our analysis.
+The university data is only available in PDF format to download. To process the data, it had
+to be transformed into a more versatile format, using two steps. First, it has been transformed into CSV, which resulted in a corrupted file.
+A custom-written parser was then used, to restore the original data structure.
+% I am currently writing here
 
@@
     \centering
-    \includegraphics{fig/prediction_modified_10\%_increase_lags_5.pdf}
+    \includegraphics{fig/prediction_modified_10p_increase_lags_5.pdf}
     \caption{Increased salary by 10 \% in Var(5) model}

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
update

@@
 
-Our central contribution is a Vector autoregressive model (VAR) discussed in \Cref{sec:methods} that is built using just mentioned data. Combined with a
+Our central contribution is a Vector autoregressive model (VAR) discussed in \Cref{sec:methods:var} that is built using just mentioned data. Combined with a
 Partial autocorrelation function (PACF) the optimal value for lag
-is determined, see \Cref{sec:methods} \textcolor{red}{subsection}, to ensure that relevant structure in the data is captured by the model.
+is determined, see \Cref{sec:methods:pacf}, to ensure that relevant structure in the data is captured by the model.
 After verifying that the VAR model functions as expected and testing the model on a test set, \Cref{sec:methods} \textcolor{red}{subsection}, we
@@
 
+\subsection{Data Processing}\label{sec:methods:processing}
 Each data set used in the analysis had to be processed, to work for our analysis.
@@
 A custom-written parser was then used, to restore the original data structure.
-% I am currently writing here
+In the year 2012, two years of high school graduates finished, due to the transition from G9 to G8 in
+Baden-Württemberg. This outlier in the data had to be flattened out to ensure the VAR model could be fitted correctly.
+The model uses the mean square error as a loss function, which makes outliers influence the data disproportionately.
+Half the number of students from the year 2012 has been spread over the next four years to flatten the data.
+The Deutsches Statistisches Bundesamt from which the average gross salary is obtained, does not account for inflation in its data set.
+Since we need to maintain comparability between past years, the salary has been adjusted for a yearly cumulative inflation rate.
 
-Our main goal was to find a correlation between the salary, the graduates and the resulting amount of students enrolling a subject in the STEM field.
-For this problem we use a vector autoregressive model, which predicts a time series
-for each input time series. The model creates a linear combination for each prediction which could be expressed in a matrix-vector notation.
+\subsection{Model creation}\label{sec:methods:var}
+A vector autoregressive model is used to analyze relationships between multiple time series.
+The model creates a linear combination for each prediction which can be expressed in a matrix-vector notation.
 \begin{equation*}
@@
 \end{equation*}
-The $\epsilon$ vector approximates an error, which is induced by the model. One main benefit of the VAR model is that it also assumes that $S_{i,j}$ is also influenced by $S_{k,l}$ 
-$i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$.
-In our case, we have three input time series, the students, salary and graduates data from the Winter Semester 2007 to the Summer Semester 2021. This model has one hyperparameter which describes
-the amount of past data that should be included in our models prediction. 
+The $\epsilon$ vector approximates an error, which is induced by the model. One main benefit of the VAR model is that it also assumes
+that $S_{i,j}$ is also influenced by $S_{k,l}$ $i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$.
+A VAR model has one major advantage, over other possible approaches, which is its simplicity. Therefore its results remain interpretable.
+This model only uses one hyperparameter which describes the amount of past data that is used to make a prediction. This parameter is called the lag.
 The vector autoregressive model heavily depends on the stationarity assumption which ensures that the stochastical process stays the same over time. 
-In this context, this is assured by the following three axioms
+In this context, this is assured by the following three axioms:
 \begin{enumerate}[label=\Roman*]
@@
 \end{enumerate}
-We assume that our time series is mostly stationary, and this is enough for our case because we are only predicting 2 data points in the future.
-In order to find the optimal hyperparameter we used two well-established methods, the PACF and also an information-based criteria particularly the AIC and 
-BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ onto $S_{n}$, and then it builds a $(1-\alpha)$ confidence interval. We choose $\alpha = 5\%$ and get
-get a 95\% Confidence Interval 
+Since we are only interested in predicting two time steps into the future, which is equivalent to one year,
+we can safely assume that our data is stationary.
+
+\subsection{PACF}\label{sec:methods:pacf}
+To find the optimal hyperparameter we used two well-established methods, the PACF and information-based criteria. Particularly the AIC and 
+BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ on $S_{n}$ to build a $(1-\alpha)$ confidence interval.
+We choose $\alpha = 5\%$ and get get a 95\% Confidence Interval 
 \begin{table}[H]
@@
         \hline
-        Prediciton 1 & 1.25723662 & 1.27163776 & 3.94417629 \\
+        Prediciton 1 & 1.257\% & 1.271\% & 3.944\% \\
         \hline
-        Prediciton 2 & 4.84256257 & 0.11922267 & 1.72230067
+        Prediciton 2 & 4.843\% & 0.119\% & 1.722\%
     \end{tabular}
-    \caption{Percentage difference between true value and predicted value for students, salary and graduates for lag 1}
+    \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to 1.}
 \end{table}
-As you can see, the percent difference between the true values especially for the students are quite high, becasue of that we tried the AIC and BIC to find a better lag for the 
-hyperparameter p.
-The abbreviation AIC stands for Akaike Information Criterion and tries to find an optimum between the complexity of the model and fitting the data, which should pretend overfitting.
-BIC is an abbreviation for Bayesian Information Criterion and the main difference between AIC and BIC is that BIC penalizes more complex models, which overall leads to less complicated models than the ones from AIC.
-The optimal lag based on AIC is 6, while for BIC it is 1, which aligns with the PACF.
+The percentage differences between the true values especially for the students are quite high. To get better results
+the AIC and BIC are used to find a better lag for the hyperparameter $p$ \textcolor{red}{is $p$ = lag?}. The abbreviation AIC
+stands for Akaike Information Criterion and tries to find an optimum between the complexity of the model and fitting the data.
+This should prevent overfitting. 
+BIC is short for Bayesian Information Criterion and penalizes more complex models.
+Therefore BIC leads to less complicated models than the ones from AIC, which sets the two apart.
+The optimal lag based on AIC is 6, while for BIC it is 1, which is consistent with the PACF.
 \begin{table}[H]
@@
         \hline
-        Prediciton 1 & 0.48413863\% & 0.41885829\% & 0.11117345\% \\
+        Prediciton 1 & 0.484\% & 0.419\% & 0.111\% \\
         \hline
-        Prediciton 2 & 1.34596809\% & 0.30583828\% & 0.93859448\%
+        Prediciton 2 & 1.346\% & 0.306\% & 0.939\%
     \end{tabular}
-    \caption{Percentage difference between true value and predicted value for students, salary and graduates for lag 5}
+    \caption{Percentage difference between the true value and predicted value for students, salary, and graduates. Lag set to 5.}
 \end{table}
+
+% I am currently here
+
 As you can see there is a major improvement in the prediction for the students in comparison to lag order 1. We decided to use lag 5, because the main focus of our 

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
update

@@
 Partial autocorrelation function (PACF) the optimal value for lag
-is determined, see \Cref{sec:methods:pacf}, to ensure that relevant structure in the data is captured by the model.
+is determined, see \Cref{sec:methods:lag}, to ensure that relevant structure in the data is captured by the model.
 After verifying that the VAR model functions as expected and testing the model on a test set, \Cref{sec:methods} \textcolor{red}{subsection}, we
@@
 
-\subsection{PACF}\label{sec:methods:pacf}
+\subsection{Choosing lag}\label{sec:methods:lag}
 To find the optimal hyperparameter we used two well-established methods, the PACF and information-based criteria. Particularly the AIC and 
@@
 
+The percentage errors show that there is a significant improvement in the prediction of student numbers compared to the lag order of 1.
+Setting lag to 5 results in even better errors for the domains of our interest, predicting students.
+
+\subsection{Test for causality}\label{sec:methods:causality}
+To ensure that the model does not try to find structure in the data, where there is none, we applied the Granger causality test.
+Granger causal implies that one time series is the result of another shifted time series.
+In this case, we try to find out if e.g. the salary time series is just a shifted version of the student's or the graduate's time series.
+The test does verifies this by using series $S_i$ and trying to insert data points from $S_j$ or $S_k$ $i \neq j\neq k$
+to predict the outcome of the model. After inserting the data points a t-test and a F-test are performed. If both
+tests return statistical significance, then we showed that $S_i$ could be expressed by a shifted input time series $S_j$ with $j\neq i$.
+\textcolor{red}{OK, and did we show that?}
+
 % I am currently here
-
-As you can see there is a major improvement in the prediction for the students in comparison to lag order 1. We decided to use lag 5, because the main focus of our 
-of our research is to predict the students based on salary and graduates.
-Another quite important aspect in the correlation analysis is the Granger Causality Test which we did on the output. 
-Granger Causal implies that one time series is the result of another shifted time series.
-In our case, we try to find out if e.g. the salary time series is just a shifted version of the students or the graduates time series.
-The Granger causality test does exactly this, it uses the time series $S_i$ and tries to insert data points from $S_j$ or $S_k$ $i \neq j\neq k$
-to predict the outcome of the model. After inserting the data points the test performs a t-test and a F-test, if both of these tests return statistically significant,
-then we showed that $S_i$ could be expressed by a shifted input time series $S_j$ with $j\neq i$
 

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
added conclusion

@@
 \section{Discussion \& Conclusion}\label{sec:conclusion}
+In our research, we used a vector autoregressive model to predict the amount of students in the next semester. 
+We used the salary data from the past 15 years and data about high school graduates in Baden-Würrtemberg as a prior. We optimized our Model to the best possible capacity,
+but there are strict limitations because vector autoregressive models are only able to detect linear patterns in the data. This occurs, because we only have linear combinations,
+based on values $S_{i,j}$ and coefficients to scale those values. If we wanted to improve our prediction we have to replace our model with other machine learning applications.
+A possible option for this would be Random Forests.\\
+Our model is a great way to show insight information about the relation of students, graduates and the expected average salary after their university degree.
+Unlike our main assumption, that the salary expectation is the striving factor for students to go into the STEM Sector it just shows a small correlation,
+between the predicted students and the salary. It seems like the main factor for a change in the students is mainly influenced by the amount of graduating high school students.
+This insight gives us hope, that most of the STEM students do what they love not for the sake of money.
 
 \section*{Contribution Statement}
-
 Explain here, in one sentence per person, what each group member contributed.

### 2024-01-28 doc/DataLiteracyStudentProject/paper.tex
visualise coefficients

@@
 
+\begin{figure*}
+    \includegraphics{fig/prediction_5.pdf}
+    \caption{VAR Model Prediction with optimal lag value of 5 on the test data set.}
+\end{figure*}
+
 \subsection{Data Processing}\label{sec:methods:processing}
@@
 \begin{equation*}
-    \resizebox{\linewidth}{!}{%
+    \resizebox{\linewidth}{!}{
         $\begin{bmatrix}
@@
 was the lowest for the students.
-\begin{figure*}
-    \includegraphics{fig/prediction_5.pdf}
-    \caption{VAR Model Prediction with hyperparameter p=5}
-\end{figure*}
+
+\begin{figure}
+    \includegraphics{fig/params_Students.pdf}
+    \caption{Coefficients of VAR model to predict students time-series
+    for $t$. $t_1$ to $t_5$ are the values of each time series respectively at the lagged time step.
+    $t_1$ is the most recent value.}
+\end{figure}
 During our analysis, we found out, that the salary has a major influence on our prediction. If our model works correctly, then the amount
@@
     \includegraphics{fig/prediction_modified_10p_increase_lags_5.pdf}
-    \caption{Increased salary by 10 \% in Var(5) model}
+    \caption{Increased salary by 10\%}
 \end{figure}

### 2024-01-28 src/salaryExpHelpers.py

@@ -12,11 +12,12 @@ def plot(data, forecast, years, name, lags) -> plt.Figure:
     dataWithForecast = np.vstack((data, forecast))
     yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
 
-    plt.rcParams.update(bundles.icml2022(column="full", nrows=1, ncols=1, usetex=False))
+    plt.rcParams.update(bundles.icml2022(column="half", nrows=1, ncols=1, usetex=False))
     fig, ax1 = plt.subplots()
     plt.title('University Data Over the Years')
     plt.xticks(rotation=30)
 
+    ax1.set_ylim([0, 10000])
     ax1.set_ylabel('Total students', color=COL_STUDENT)
     # Plot main data
     ax1.bar(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 0], color=COL_STUDENT, label='Enrolled students')
@@ -29,11 +30,11 @@ def plot(data, forecast, years, name, lags) -> plt.Figure:
     ax2.set_ylabel('Sallary', color=COL_SALARY)
 
     # Plot main data
-    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average salary after tax')
+    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average salary gross')
 
     # Plot forecast with a different color
     ax2.plot(yearsWithForecast[-len(forecast) - 1:], dataWithForecast[-len(forecast) - 1:, 1],
-             color=COL_SALARY_PRED, label='Average salary after tax forecast')
+             color=COL_SALARY_PRED, label='Average salary gross forecast')
 
     # Create one legend for both subplots
     lines, labels = ax1.get_legend_handles_labels()

### 2024-01-28 src/salaryExpHelpers.py
change plot titles and captions

@@ -14,7 +14,7 @@ def plot(data, forecast, years, name, lags) -> plt.Figure:
 
     plt.rcParams.update(bundles.icml2022(column="half", nrows=1, ncols=1, usetex=False))
     fig, ax1 = plt.subplots()
-    plt.title('University Data Over the Years')
+    plt.title('Enrolled students with altered salary')
     plt.xticks(rotation=30)
 
     ax1.set_ylim([0, 10000])

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
change plot titles and captions
finished Results part

@@
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
+\newcommand{\figref}[1]{Fig.~\ref{#1}}
 
@@
     \includegraphics{fig/prediction_5.pdf}
-    \caption{VAR Model Prediction with optimal lag value of 5 on the test data set.}
+    \caption{VAR model Prediction with optimal lag value of 5 on the test data set.}
 \end{figure*}
@@
 
+\subsection{Fitted model}\label{sec:methods:fit}
+\begin{figure}
+    \includegraphics{fig/params_Students.pdf}
+    \caption{Coefficients of VAR model to predict students time-series
+    for $t$. $t_1$ to $t_5$ are the values of each time series respectively at the lagged time step.
+    $t_1$ is the most recent value.}
+\end{figure}
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{c|c|c|c}
+        & students & salary & graduates\\
+        \hline
+        students & 1.000000 & -0.115442 &  0.111069\\
+        \hline
+        salary & -0.115442 & 1.000000 & -0.902321\\
+        \hline
+        graduates & 0.111069 & -0.902321 &  1.000000\\
+    \end{tabular}
+    \caption{Correlation Matrix of residuals using 5 lags.}
+\end{table}
+
 % I am currently here
+% Here the explanation for the correlation matrix
+% Here teh the explanation for the visualisation of coefficients
 
@@
 \section{Results}\label{sec:results}
-The hyperparameter optimization is always a trade-off between several optimal values for different parameters. 
-Our main goal was to predict the students who enroll in study subjects in the STEM sector when using the salary and the
-graduates of the given years as prior knowledge. For this reason, we used a lag of 5, because the overall percentage error
-was the lowest for the students.
+It is not possible to find a lag value that gives the best results for all three time series.
+However one can set the optimal lag value such that it suits forecasting students best (\Cref{sec:methods:lag}).
+With the obtained insights into the relation between the number of enrolled students, salary expectations, and number of graduates,
+it is now possible to make predictions on different scenarios.
 
-\begin{figure}
-    \includegraphics{fig/params_Students.pdf}
-    \caption{Coefficients of VAR model to predict students time-series
-    for $t$. $t_1$ to $t_5$ are the values of each time series respectively at the lagged time step.
-    $t_1$ is the most recent value.}
-\end{figure}
-During our analysis, we found out, that the salary has a major influence on our prediction. If our model works correctly, then the amount
-of predicted students should go down, because fewer students imply that there is a lack of qualified workers on the job market.
-This on the other hand leads to an increase in the salary. Now we manipulated the salary data with a fixed $\alpha\in \mathbb{N}$ to increase
-the whole time series by a constant factor. We used this approach because the vector autoregressive model internally uses Mean Squared Error as loss function.
-The Mean Squared Error squares the difference between the data points, which leads to a stronger influence of large differences.
-This high error leads to a wrong minimization, which gives wrong coefficients.
-Another more important point is that we don't want to violate stationarity, which ensures that our model predicts in a good manner.\\
-\begin{table}[H]
-    \centering
-    \begin{tabular}{c|c|c|c}
-        & students & salary & graduates\\
-        \hline
-        students & 1.000000 & -0.115442 &  0.111069\\
-        \hline
-        salary & -0.115442 & 1.000000 & -0.902321\\
-        \hline
-        graduates & 0.111069 & -0.902321 &  1.000000\\
-    \end{tabular}
-    \caption{Correlation Matrix of Var(5) model}
-\end{table}
+The training of the model stays set, and the whole dataset available is used. It is now possible to 
+generate a new data set with a 10\% increase in salary expectations and set it as the new model input, \figref{fig:altered salary}.
+The observed change in the prediction supports the obtained results in \Cref{sec:methods:fit}. There is a significant
+dependence between salary expectations and the number of students enrolled in STEM courses. As soon as the salary
+increases, there is also an upward trend visible in student numbers. The observed change is not massive which can be explained
+by the fact that the mass of students are students in higher semesters. The only direct major influence on
+the total number of enrolled students is made by people deciding to start this year. There won't be a lot of people changing subjects
+halfway through their academic careers.
 
-As you can see the correlation matrix, tells us that the model also takes the relationship into account, that a 
-manipulated salary has an inverse relationship to the amount of students.
+This also explains the huge spike in predicting the average gross salary. In the first few semesters,
+the amount of new students is not high enough, to capture the need for new people. Therefore the model suggests raising the
+salary even higher, to accommodate for that.
 
@@
     \includegraphics{fig/prediction_modified_10p_increase_lags_5.pdf}
-    \caption{Increased salary by 10\%}
+    \caption{Predict the number of students enrolled in STEM courses using an increased salary by 10\%.}
+    \label{fig:altered salary}
 \end{figure}
@@
 A possible option for this would be Random Forests.\\
-Our model is a great way to show insight information about the relation of students, graduates and the expected average salary after their university degree.
+Our model is a great way to show insight information about the relation of students, graduates, and the expected average salary after their university degree.
 Unlike our main assumption, that the salary expectation is the striving factor for students to go into the STEM Sector it just shows a small correlation,
@@
 \section*{Contribution Statement}
-Explain here, in one sentence per person, what each group member contributed.
-For example, you could write: Max Mustermann collected and prepared data.
+\textcolor{red}{TODO}
+Abdallah Abdul-Latif analyzed the optimal parameter lag and mapped courses to salary sectors.
+Lisa-Maria Fritsch
+Paul Kaifler 
+Maximilian Schnitt parsed university data and analyzed optimal parameter lag.
+All authors jointly did the analyses of the data and wrote the text of the report.
+% Max Mustermann collected and prepared data.
 Gabi Musterfrau and John Doe performed the data analysis.
@@
 
-\section*{Notes}
+% \section*{Notes}
 
-Your entire report has a \textbf{hard page limit of 4 pages} excluding references.
+% Your entire report has a \textbf{hard page limit of 4 pages} excluding references.
 (I.e.

### 2024-01-29 src/salaryExpHelpers.py

@@ -8,8 +8,9 @@ from src.constants import COL_STUDENT, COL_STUDENT_PRED, COL_SALARY, COL_SALARY_
 from tueplots import bundles
 
 
-def plot(data, forecast, years, name, lags) -> plt.Figure:
+def plot(data, forecast, ogForecast, years, name, lags) -> plt.Figure:
     dataWithForecast = np.vstack((data, forecast))
+    ogDataWithForecast = np.vstack((data, ogForecast))
     yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
 
     plt.rcParams.update(bundles.icml2022(column="half", nrows=1, ncols=1, usetex=False))
@@ -26,15 +27,19 @@ def plot(data, forecast, years, name, lags) -> plt.Figure:
     ax1.bar(yearsWithForecast[-len(forecast):], dataWithForecast[-len(forecast):, 0],
             color=COL_STUDENT_PRED, label='Enrolled students forecast')
 
+    # Plot original forecast with a different color
+    ax1.bar(yearsWithForecast[-len(forecast):], ogDataWithForecast[-len(forecast):, 0],
+            color='silver', label='Original prediction', alpha=0.5)
+
     ax2 = ax1.twinx()  # instantiate a second axes that shares the same x-axis
     ax2.set_ylabel('Sallary', color=COL_SALARY)
 
     # Plot main data
-    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average salary gross')
+    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average gross salary')
 
     # Plot forecast with a different color
     ax2.plot(yearsWithForecast[-len(forecast) - 1:], dataWithForecast[-len(forecast) - 1:, 1],
-             color=COL_SALARY_PRED, label='Average salary gross forecast')
+             color=COL_SALARY_PRED, label='Average gross salary  forecast')
 
     # Create one legend for both subplots
     lines, labels = ax1.get_legend_handles_labels()

### 2024-01-29 src/salaryExpHelpers.py
Paper/coef (#39)
* visualise coefficients

* change plot titles and captions

* finished Results part

* Update plot title

@@ -8,15 +8,17 @@ from src.constants import COL_STUDENT, COL_STUDENT_PRED, COL_SALARY, COL_SALARY_
 from tueplots import bundles
 
 
-def plot(data, forecast, years, name, lags) -> plt.Figure:
+def plot(data, forecast, ogForecast, years, name, lags) -> plt.Figure:
     dataWithForecast = np.vstack((data, forecast))
+    ogDataWithForecast = np.vstack((data, ogForecast))
     yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
 
-    plt.rcParams.update(bundles.icml2022(column="full", nrows=1, ncols=1, usetex=False))
+    plt.rcParams.update(bundles.icml2022(column="half", nrows=1, ncols=1, usetex=False))
     fig, ax1 = plt.subplots()
-    plt.title('University Data Over the Years')
+    plt.title('Enrolled students with altered salary')
     plt.xticks(rotation=30)
 
+    ax1.set_ylim([0, 10000])
     ax1.set_ylabel('Total students', color=COL_STUDENT)
     # Plot main data
     ax1.bar(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 0], color=COL_STUDENT, label='Enrolled students')
@@ -25,15 +27,19 @@ def plot(data, forecast, years, name, lags) -> plt.Figure:
     ax1.bar(yearsWithForecast[-len(forecast):], dataWithForecast[-len(forecast):, 0],
             color=COL_STUDENT_PRED, label='Enrolled students forecast')
 
+    # Plot original forecast with a different color
+    ax1.bar(yearsWithForecast[-len(forecast):], ogDataWithForecast[-len(forecast):, 0],
+            color='silver', label='Original prediction', alpha=0.5)
+
     ax2 = ax1.twinx()  # instantiate a second axes that shares the same x-axis
     ax2.set_ylabel('Sallary', color=COL_SALARY)
 
     # Plot main data
-    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average salary after tax')
+    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average gross salary')
 
     # Plot forecast with a different color
     ax2.plot(yearsWithForecast[-len(forecast) - 1:], dataWithForecast[-len(forecast) - 1:, 1],
[…]

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
Added Fitted Model section

@@
 \newcommand{\theHalgorithm}{\arabic{algorithm}}
-\newcommand{\figref}[1]{Fig.~\ref{#1}}
 
@@
 \section{Data and Methods}\label{sec:methods}
-% Mabe focus here on what we did with the data
-% I.e. inflation adjustment
-% flattening G8 G9 conflict
 
@@
     \caption{Coefficients of VAR model to predict students time-series
-    for $t$. $t_1$ to $t_5$ are the values of each time series respectively at the lagged time step.
-    $t_1$ is the most recent value.}
+    for $t_0$. $t_{-1}$ to $t_{-5}$ are the values of each time series respectively at the lagged time step.
+    $t_{-1}$ is the most recent value.}
+    \label{fig:student coefficients}
 \end{figure}
+
+% I am currently here
+% Here the explanation for the correlation matrix
+% Here teh the explanation for the visualisation of coefficients
+
+With the correct lag chosen, the model fits the data by minimizing the mean square error.
+The returned coefficients describe the impact of a specific lagged value from one time series
+on predicting the next value. \ref{fig:student coefficients} visualizes all coefficients
+relevant for predicting the number of students in the next semester. 
+Notably, salary is the most relevant information for the prediction. All 5
+coefficient values for each time step are non-zero, indicating a strong correlation.
+Graduates, on the other hand, are not as relevant, recognizable by values close to zero.
+
+The suggested lag value of $5$, earlier obtained in \Cref{sec:methods:lag} is retrievable as well.
+The absolute value of the coefficient for predicting students on $t_{-5}$ is the third largest, adding
+a lot of information to the model.
+
+To further understand the data underlying correlations, one can look at the correlation matrix
+of residuals in table \ref{table:correlation matrix}.
+This matrix describes the degree to which the model's errors, called residuals, are related to each other.
+High correlations between errors can indicate a not fully captured underlying structure.
+Values close to zero, like $-0.115$ between salary and graduates, indicate that the errors
+in predicting the salary variable are not related to the errors in predicting the graduates variable.
 
@@
         \hline
-        students & 1.000000 & -0.115442 &  0.111069\\
+        students & 1.000 & -0.115 &  0.111\\
         \hline
-        salary & -0.115442 & 1.000000 & -0.902321\\
+        salary & -0.115 & 1.000000 & -0.902\\
         \hline
-        graduates & 0.111069 & -0.902321 &  1.000000\\
+        graduates & 0.111 & -0.902 &  1.000\\
     \end{tabular}
-    \caption{Correlation Matrix of residuals using 5 lags.}
+    \caption{Correlation matrix of residuals using 5 lags.}
+    \label{table:correlation matrix}
 \end{table}
-
-% I am currently here
-% Here the explanation for the correlation matrix
-% Here teh the explanation for the visualisation of coefficients
 
@@
 The training of the model stays set, and the whole dataset available is used. It is now possible to 
-generate a new data set with a 10\% increase in salary expectations and set it as the new model input, \figref{fig:altered salary}.
+generate a new data set with a 10\% increase in salary expectations and set it as the new model input, figure \ref{fig:altered salary}.
 The observed change in the prediction supports the obtained results in \Cref{sec:methods:fit}. There is a significant

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
latex formatting

@@
     \includegraphics{fig/prediction_5.pdf}
-    \caption{VAR model Prediction with optimal lag value of 5 on the test data set.}
+    \caption{VAR model Prediction with optimal lag value of $5$ on the test data set.}
 \end{figure*}
@@
     \end{tabular}
-    \caption{Percentage difference between the true value and predicted value for students, salary, and graduates. Lag set to 5.}
+    \caption{Percentage difference between the true value and predicted value for students, salary, and graduates. Lag set to $5$.}
 \end{table}
@@
 The percentage errors show that there is a significant improvement in the prediction of student numbers compared to the lag order of 1.
-Setting lag to 5 results in even better errors for the domains of our interest, predicting students.
+Setting lag to $5$ results in even better errors for the domains of our interest, predicting students.
 
@@
 
-% I am currently here
-% Here the explanation for the correlation matrix
-% Here teh the explanation for the visualisation of coefficients
-
 With the correct lag chosen, the model fits the data by minimizing the mean square error.
@@
 relevant for predicting the number of students in the next semester. 
-Notably, salary is the most relevant information for the prediction. All 5
+Notably, salary is the most relevant information for the prediction. All five
 coefficient values for each time step are non-zero, indicating a strong correlation.
@@
     \end{tabular}
-    \caption{Correlation matrix of residuals using 5 lags.}
+    \caption{Correlation matrix of residuals using $5$ lags.}
     \label{table:correlation matrix}

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
Updated Discussions

@@
 We used the salary data from the past 15 years and data about high school graduates in Baden-Würrtemberg as a prior. We optimized our Model to the best possible capacity,
-but there are strict limitations because vector autoregressive models are only able to detect linear patterns in the data. This occurs, because we only have linear combinations,
-based on values $S_{i,j}$ and coefficients to scale those values. If we wanted to improve our prediction we have to replace our model with other machine learning applications.
-A possible option for this would be Random Forests.\\
-Our model is a great way to show insight information about the relation of students, graduates, and the expected average salary after their university degree.
-Unlike our main assumption, that the salary expectation is the striving factor for students to go into the STEM Sector it just shows a small correlation,
-between the predicted students and the salary. It seems like the main factor for a change in the students is mainly influenced by the amount of graduating high school students.
-This insight gives us hope, that most of the STEM students do what they love not for the sake of money.
+but there are strict limitations that can not be overcome. Vector autoregressive models are only capable of detecting linear patterns in the data.
+This occurs, because we only have linear combinations, based on values $S_{i,j}$ and coefficients to scale those values. If we wanted to improve
+our prediction we had to replace our model with other machine learning applications.
+A promising approach for this could be to implement a Random Forest algorithm.\\
+Our model on the other hand compensates for any potential lag in accuracy by being easily interpretable. Each parameter of the model influences exactly
+one part of the predictions and is therefore transparent in its process.
+Also worth mentioning is that our predictions are only based on three different inputs. There are certainly other 
+variables impacting the number of students enrolled in courses at a university. Especially ones, that are not easily quantifiable and are subject to
+personal preferences. However, we are confident that our selected set of data captures the main aspects, as our model performs great on the provided test set.
 

### 2024-01-29 src/constants.py
reformatted all notebooks

@@ -12,8 +12,8 @@ STUDENTS_PATH = os.path.join(DATA_DIR, "student_data_per_subject.xlsx")
 INFLATION_PATH = os.path.join(DATA_DIR, "inflation_germany.csv")
 EXP_DIR = os.path.join(PROJECT_DIR, "exp")
 DOC_PATH = os.path.join(PROJECT_DIR, "doc", "DataLiteracyStudentProject")
-
 FIG_DIR = os.path.join(DOC_PATH, "fig")
+
 # paths for the paper
 PAPER_DIR = os.path.join(PROJECT_DIR, "doc/DataLiteracyStudentProject")
 TEX_PATH = os.path.join(PAPER_DIR, "paper.tex")
@@ -82,6 +82,7 @@ YEARS = [
 ]
 
 COURSES = ['"Biologie"', '"Chemie"', '"Geowissenschaften"', '"Informatik"', '"Mathematik"', '"Pharmazie"', '"Biochemie"', '"Physik"', '"Psychologie"']
+
 SECTORS = ['Herstellung von pharmazeutischen Erzeugnissen',
     'Herstellung von chemischen Erzeugnissen',
     'Forschung und Entwicklung',

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex

@@
 It contains the inflation rate in Germany for each year.
-The number of students is also limited by a further parameter. Namely the quantity of people eligible for an academic career.
+The number of students is also limited by a further parameter. Namely the number of people eligible for an academic career.
 The \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} provides
@@
 Our central contribution is a Vector autoregressive model (VAR) discussed in \Cref{sec:methods:var} that is built using just mentioned data. Combined with a
-Partial autocorrelation function (PACF) the optimal value for lag
+Partial autocorrelation function (PACF) the optimal value for the lag
 is determined, see \Cref{sec:methods:lag}, to ensure that relevant structure in the data is captured by the model.

### 2024-01-29 src/salaryExpHelpers.py
removed typos

@@ -47,7 +47,7 @@ def plot(data: np.ndarray, forecast: np.ndarray, ogForecast: np.ndarray, years: 
                 color='silver', label='Original prediction', alpha=0.5)
 
         ax2 = ax1.twinx()  # instantiate a second axes that shares the same x-axis
-        ax2.set_ylabel('Sallary', color=COL_SALARY)
+        ax2.set_ylabel('Salary', color=COL_SALARY)
 
         # Plot main data
         ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average gross salary')

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
small changes

@@
 
-The analysis relies on three underlying datasets.
+The analysis is built on three underlying datasets.
 The first dataset captures the historical enrollment figures of students at the University of Tübingen. 
@@
 Our central contribution is a Vector autoregressive model (VAR) discussed in \Cref{sec:methods:var} that is built using just mentioned data. Combined with a
-Partial autocorrelation function (PACF) the optimal value for the lag
+partial autocorrelation function (PACF) the optimal value for the lag
 is determined, see \Cref{sec:methods:lag}, to ensure that relevant structure in the data is captured by the model.
-After verifying that the VAR model functions as expected and testing the model on a test set, \Cref{sec:methods} \textcolor{red}{subsection}, we
+After verifying that the VAR model functions as expected and testing the model on a test set, \Cref{fig:accuracy model} \textcolor{red}{subsection}, we
 can predict all three input time series. As we are only interested in the number of students enrolled and the development of the salary, our research is 
@@
     \caption{VAR model Prediction with optimal lag value of $5$ on the test data set.}
+    \label{fig:accuracy model}
 \end{figure*}
@@
     \end{tabular}
-    \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to 1.}
+    \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to $1$.}
 \end{table}
@@
     \end{tabular}
-    \caption{Percentage difference between the true value and predicted value for students, salary, and graduates. Lag set to $5$.}
+    \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to $5$.}
 \end{table}
@@
 The returned coefficients describe the impact of a specific lagged value from one time series
-on predicting the next value. \ref{fig:student coefficients} visualizes all coefficients
+on predicting the next value. \Cref{fig:student coefficients} visualizes all coefficients
 relevant for predicting the number of students in the next semester. 
@@
 To further understand the data underlying correlations, one can look at the correlation matrix
-of residuals in table \ref{table:correlation matrix}.
+of residuals in \Cref{table:correlation matrix}.
 This matrix describes the degree to which the model's errors, called residuals, are related to each other.
@@
 The training of the model stays set, and the whole dataset available is used. It is now possible to 
-generate a new data set with a 10\% increase in salary expectations and set it as the new model input, figure \ref{fig:altered salary}.
+generate a new data set with a 10\% increase in salary expectations and set it as the new model input, \Cref{fig:altered salary}.
 The observed change in the prediction supports the obtained results in \Cref{sec:methods:fit}. There is a significant

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
Small adjustments

@@
 It contains the salary trend for all major industrial sectors in Germany. This dataset does not adjust for inflation,
-therefore an additional \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=61121-0001&bypass=true&levelindex=0&levelid=1706446467491#abreadcrumb}{dataset} from the Deutsches Statistisches Bundesamt is used to account for that.
-It contains the inflation rate in Germany for each year.
+therefore an additional \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=61121-0001&bypass=true&levelindex=0&levelid=1706446467491#abreadcrumb}{dataset}
+from the Deutsches Statistisches Bundesamt is used to account for that (\Cref{sec:methods:processing}). It contains the inflation rate in Germany for each year.
 The number of students is also limited by a further parameter. Namely the number of people eligible for an academic career.
@@
 is determined, see \Cref{sec:methods:lag}, to ensure that relevant structure in the data is captured by the model.
-After verifying that the VAR model functions as expected and testing the model on a test set, \Cref{fig:accuracy model} \textcolor{red}{subsection}, we
+After verifying that the VAR model functions as expected and testing the model on a test set, \Cref{fig:accuracy model}, we
 can predict all three input time series. As we are only interested in the number of students enrolled and the development of the salary, our research is 
 focused on these two instances.
-This allows us to conclude that salary has a significant impact on the amount of STEM students. Further, we can model different salary development scenarios, 
+This allows us to conclude in \Cref{sec:methods:causality} that salary has a significant impact on the amount of STEM students. Further, we can model different salary development scenarios, 
 and see how the students would likely react (\Cref{sec:results}).
@@
     \includegraphics{fig/prediction_5.pdf}
-    \caption{VAR model Prediction with optimal lag value of $5$ on the test data set.}
+    \caption{VAR model Prediction with optimal lag value of $5$ on the test data set. Standard error with confidence interval value $\alpha = 0.05$.}
     \label{fig:accuracy model}

### 2024-01-29 src/constants.py
Added documentation

@@ -1,4 +1,4 @@
-# Description: Constants used in the code
+"""Project constants"""
 import os
 import re
 from tueplots.constants.color import rgb
@@ -81,16 +81,24 @@ YEARS = [
     "WiSe 2023/2024",
 ]
 
-COURSES = ['"Biologie"', '"Chemie"', '"Geowissenschaften"', '"Informatik"', '"Mathematik"', '"Pharmazie"', '"Biochemie"', '"Physik"', '"Psychologie"']
+COURSES = [
+    '"Biologie"',
+    '"Chemie"',
+    '"Geowissenschaften"',
+    '"Informatik"',
+    '"Mathematik"',
+    '"Pharmazie"',
+    '"Biochemie"',
+    '"Physik"',
+    '"Psychologie"']
 
 SECTORS = ['Herstellung von pharmazeutischen Erzeugnissen',
-    'Herstellung von chemischen Erzeugnissen',
-    'Forschung und Entwicklung',
-    'Gewinnung von Steinen und Erden, sonstiger Bergbau',
-    'Dienstleistungen der Informationstechnologie',
-    'H.v. DV-Geräten, elektron. u. opt. Erzeugnissen',
-    'Wirtschaftsprüfung und Steuerberatung, Buchführung',
-    'H.v. Bestrahlungs- u.a. elektromed. Geräten',
-    'Gesundheitswesen',
-]
-
+           'Herstellung von chemischen Erzeugnissen',
+           'Forschung und Entwicklung',
+           'Gewinnung von Steinen und Erden, sonstiger Bergbau',
+           'Dienstleistungen der Informationstechnologie',
+           'H.v. DV-Geräten, elektron. u. opt. Erzeugnissen',
+           'Wirtschaftsprüfung und Steuerberatung, Buchführung',
+           'H.v. Bestrahlungs- u.a. elektromed. Geräten',
+           'Gesundheitswesen',
+           ]

### 2024-01-29 src/dataCheck.py

@@ -1,3 +1,4 @@
+"""This is a small test file to check if the data is parsed correctly"""
 import unittest
 from constants import STUDENTS_PATH, SEMESTERS
 import pandas as pd

### 2024-01-29 src/dataParser.py

@@ -1,3 +1,4 @@
+"""This module parses the data from the csv files and saves as it as an excel sheet"""
 import os
 import re
 import pandas as pd

### 2024-01-29 src/salaryExpHelpers.py

@@ -1,3 +1,4 @@
+"""Helper functions for salary augmentation experiments"""
 import matplotlib.pyplot as plt
 
 import copy
@@ -11,9 +12,9 @@ from tueplots import bundles
 
 
 def plot(data: np.ndarray, forecast: np.ndarray, ogForecast: np.ndarray, years: List[str], name: str, lags: int) -> plt.Figure:
-        """ creates plot, that visualizes the influence of the manipulated data
+    """ creates plot, that visualizes the influence of the manipulated data
 
-        Args:
+    Args:
             data (: np.ndarray): the completeData used for training the model
             forecast (: np.ndarray): prediction from model with given data
             ogForecast (: np.ndarray): original prediction, without data manipulation
@@ -21,62 +22,62 @@ def plot(data: np.ndarray, forecast: np.ndarray, ogForecast: np.ndarray, years: 
             name (str): filename for saving
             lags (int): the lags we consider
 
-        Returns:
-            plt.Figure: creates matplotlib plot that shows difference between 
-        """
-        dataWithForecast = np.vstack((data, forecast))
-        ogDataWithForecast = np.vstack((data, ogForecast))
-        yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
+    Returns:
+            plt.Figure: creates matplotlib plot that shows difference between
+    """
+    dataWithForecast = np.vstack((data, forecast))
+    ogDataWithForecast = np.vstack((data, ogForecast))
+    yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
 
-        plt.rcParams.update(bundles.icml2022(column="half", nrows=1, ncols=1, usetex=False))
-        fig, ax1 = plt.subplots()
-        plt.title('Enrolled students with altered salary')
-        plt.xticks(rotation=30)
+    plt.rcParams.update(bundles.icml2022(column="half", nrows=1, ncols=1, usetex=False))
+    fig, ax1 = plt.subplots()
+    plt.title('Enrolled students with altered salary')
+    plt.xticks(rotation=30)
 
-        ax1.set_ylim([0, 10000])
-        ax1.set_ylabel('Total students', color=COL_STUDENT)
-        # Plot main data
-        ax1.bar(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 0], color=COL_STUDENT, label='Enrolled students')
[…]

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
updates

@@
 \twocolumn[
-\icmltitle{All just Greedy for Money? - A prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field}
+\icmltitle{A prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field}
 
@@
 \icmlaffiliation{second}{Matrikelnummer 4189024, lisa-maria.fritsch@student.uni-tuebingen.de, MSc Computer Science}
-\icmlaffiliation{third}{Matrikelnummer 12345678, first.last@student.uni-tuebingen.de, MSc Media Informatics}
+\icmlaffiliation{third}{Matrikelnummer 5993286, paul.kaifler@student.uni-tuebingen.de, Bsc Computer Science}
 \icmlaffiliation{fourth}{Matrikelnummer 6040570, maximilian.schnitt@student.uni-tuebingen.de, BSc Computer Science}
@@
 
+A repository containing the complete analyses and results can be found \href{https://github.com/mxs01/AnalysingStudentDevelopment}{here}.
+
 % In this section, describe \emph{what you did}. Roughly speaking, 
@@
 The $\epsilon$ vector approximates an error, which is induced by the model. One main benefit of the VAR model is that it also assumes
-that $S_{i,j}$ is also influenced by $S_{k,l}$ $i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$.
+that $S_{i,j}$ is also influenced by $S_{k,l}$ $i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$. As mentioned in \citep{kot}
 A VAR model has one major advantage, over other possible approaches, which is its simplicity. Therefore its results remain interpretable.
@@
 BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ on $S_{n}$ to build a $(1-\alpha)$ confidence interval.
-We choose $\alpha = 5\%$ and get get a 95\% Confidence Interval 
+We choose $\alpha = 5\%$ and get a 95\% Confidence Interval. The optimal value according to PACF is $3$.
 \begin{table}[H]
@@
         \hline
-        Prediciton 1 & 1.257\% & 1.271\% & 3.944\% \\
+        Prediciton 1 & 0.704\% & 0.871\% & 2.658\% \\
         \hline
-        Prediciton 2 & 4.843\% & 0.119\% & 1.722\%
+        Prediciton 2 & 2.786\% & 0.301\% & 0.286\%
     \end{tabular}
-    \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to $1$.}
+    \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to $3$.}
+    \label{tab:resid1}
 \end{table}
-The percentage differences between the true values especially for the students are quite high. To get better results
-the AIC and BIC are used to find a better lag for the hyperparameter $p$ \textcolor{red}{is $p$ = lag?}. The abbreviation AIC
+The percentage differences between the true values and the predicted ones, especially for the students, are quite high, see \Cref{tab:resid1}.
+To get better results
+the AIC and BIC are used to find a better lag for the hyperparameter $p$, which is the lag. The abbreviation AIC
 stands for Akaike Information Criterion and tries to find an optimum between the complexity of the model and fitting the data.
@@
 Therefore BIC leads to less complicated models than the ones from AIC, which sets the two apart.
-The optimal lag based on AIC is 6, while for BIC it is 1, which is consistent with the PACF.
+Both BIC and AIC return $6$ as an optimal lag value for our data set. Since we are focusing on predicting students, low errors in that domain
+are more important to us. Minimizing these errors within the range from $3$ to $6$ leads to an optimal lag value of $5$.
 \begin{table}[H]
@@
     \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to $5$.}
+    \label{tab:resid2}
 \end{table}
 
-The percentage errors show that there is a significant improvement in the prediction of student numbers compared to the lag order of 1.
-Setting lag to $5$ results in even better errors for the domains of our interest, predicting students.
+The percentage errors in \Cref{tab:resid2} show that there is a significant improvement in the prediction of student numbers
+compared to the lag order of $3$ from \Cref{tab:resid1}.
 
@@
 tests return statistical significance, then we showed that $S_i$ could be expressed by a shifted input time series $S_j$ with $j\neq i$.
-\textcolor{red}{OK, and did we show that?}
+% tbd notebook 2
+\textcolor{red}{TBD}
 
@@
     \caption{Coefficients of VAR model to predict students time-series
-    for $t_0$. $t_{-1}$ to $t_{-5}$ are the values of each time series respectively at the lagged time step.
+    for $t_0$. $t_{-1}$ to $t_{-5}$ are the values of each time series respectively at the lagged time step, which is equal to one semester.
     $t_{-1}$ is the most recent value.}
@@
 To further understand the data underlying correlations, one can look at the correlation matrix
-of residuals in \Cref{table:correlation matrix}.
+of residuals in \Cref{tab:correlation matrix}.
 This matrix describes the degree to which the model's errors, called residuals, are related to each other.
@@
     \caption{Correlation matrix of residuals using $5$ lags.}
-    \label{table:correlation matrix}
+    \label{tab:correlation matrix}
 \end{table}
@@
 \section*{Contribution Statement}
-\textcolor{red}{TODO}
 Abdallah Abdul-Latif analyzed the optimal parameter lag and mapped courses to salary sectors.
-Lisa-Maria Fritsch
-Paul Kaifler 
-Maximilian Schnitt parsed university data and analyzed optimal parameter lag.
+Lisa-Maria Fritsch was responsible for data pre-processing.
+Paul Kaifler fitted the model to the data and created visualizations.
+Maximilian Schnitt parsed the university data and analyzed optimal parameter lag.
 All authors jointly did the analyses of the data and wrote the text of the report.
@@
 More details about how to prepare the report, inclucing how to produce plots, cite correctly, and how to ideally structure your github repo, will be discussed in the lecture, where a rubric for the evaluation will also be provided.
+
+% References
 

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
Increment

@@
 \twocolumn[
-\icmltitle{A prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field}
+\icmltitle{A Prediction and Analysis of the interplay between Graduates and Salary Prospects in the STEM Field}
 
@@
 this academic career is known for being one of the most demanding.
-This paper aims to explore the potential influence of financial incentives.
+This paper aims to explore the potential influence of financial incentives, uncovering hidden relations.
 
@@
 The first dataset captures the historical enrollment figures of students at the University of Tübingen. 
-It contains the total number of students for each semester, for all study programs. This dataset is maintained and published by the
+It contains the total number of students for each semester, for all study programs ever offered. This dataset is maintained and published by the
 \href{https://uni-tuebingen.de/einrichtungen/verwaltung/iv-studierende/studierendenabteilung/statistiken/}{Statistics Bureau of the University of Tübingen}.
 The University of Tübingen makes a good candidate to discover potential structure in the number
-of enrollments because STEM programs are strongly represented here. % TODO: I think mentioning the potential risk of BIAS could be in the Discussion-part
+of enrollments because STEM programs are strongly represented here.
 To quantify the salary expectations, a dataset from the \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=62321-0001&bypass=true&levelindex=0&levelid=1702307320529#abreadcrumb}{Deutsches Statistisches Bundesamt} is used.
@@
 therefore an additional \href{https://www-genesis.destatis.de/genesis//online?operation=table&code=61121-0001&bypass=true&levelindex=0&levelid=1706446467491#abreadcrumb}{dataset}
-from the Deutsches Statistisches Bundesamt is used to account for that (\Cref{sec:methods:processing}). It contains the inflation rate in Germany for each year.
-The number of students is also limited by a further parameter. Namely the number of people eligible for an academic career.
+from the Deutsches Statistisches Bundesamt is used to account for that (\Cref{sec:methods:processing}). It contains the yearly inflation rate in Germany.
+The number of students is also limited by a further parameter. Namely the number of people eligible for an academic career, hence all high school graduates.
 The \href{https://www.statistik-bw.de/BildungKultur/SchulenAllgem/LRt0302.jsp}{Statistisches Landesamt Baden-Württemberg} provides
@@
 and see how the students would likely react (\Cref{sec:results}).
-
-A repository containing the complete analyses and results can be found \href{https://github.com/mxs01/AnalysingStudentDevelopment}{here}.
 
@@
 The $\epsilon$ vector approximates an error, which is induced by the model. One main benefit of the VAR model is that it also assumes
-that $S_{i,j}$ is also influenced by $S_{k,l}$ $i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$. As mentioned in \citep{kot}
+that $S_{i,j}$ is also influenced by $S_{k,l}$ $i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$. As mentioned in \citep{var}
 A VAR model has one major advantage, over other possible approaches, which is its simplicity. Therefore its results remain interpretable.
@@
 In this context, this is assured by the following three axioms:
+\begin{samepage}
 \begin{enumerate}[label=\Roman*]
@@
 \end{enumerate}    
+\end{samepage}
 Since we are only interested in predicting two time steps into the future, which is equivalent to one year,
@@
 \subsection{Choosing lag}\label{sec:methods:lag}
-To find the optimal hyperparameter we used two well-established methods, the PACF and information-based criteria. Particularly the AIC and 
+To find the optimal hyperparameter we used two well-established methods, the PACF and information-based criteria as descriped in \citep{pacf}. Particularly the AIC and 
 BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ on $S_{n}$ to build a $(1-\alpha)$ confidence interval.
@@
 the AIC and BIC are used to find a better lag for the hyperparameter $p$, which is the lag. The abbreviation AIC
-stands for Akaike Information Criterion and tries to find an optimum between the complexity of the model and fitting the data.
+stands for Akaike Information Criterion and it tries to find an optimum between the complexity of the model and fitting the data.
 This should prevent overfitting. 
@@
 In this case, we try to find out if e.g. the salary time series is just a shifted version of the student's or the graduate's time series.
-The test does verifies this by using series $S_i$ and trying to insert data points from $S_j$ or $S_k$ $i \neq j\neq k$
-to predict the outcome of the model. After inserting the data points a t-test and a F-test are performed. If both
-tests return statistical significance, then we showed that $S_i$ could be expressed by a shifted input time series $S_j$ with $j\neq i$.
+The test verifies this by using series $S_i$ and trying to insert data points from $S_j$ or $S_k$ $i \neq j\neq k$
+to predict the outcome of the model.
+A t-test and an F-test are performed, if both tests return statistical significance, one time series can be constructed by another.
+The Granger causality test reports back no statistical significance for our data set, which means that predicting is a nontrivial
+operation.
+
+% After inserting the data points a t-test and a F-test are performed.
+
+% If both tests return statistical significance, then we showed that $S_i$ could be expressed by a shifted input time series $S_j$ with $j\neq i$.
 % tbd notebook 2
-\textcolor{red}{TBD}
+% \textc olor{red}{TBD}
 
 \subsection{Fitted model}\label{sec:methods:fit}
-\begin{figure}
+\begin{figure}[t]
     \includegraphics{fig/params_Students.pdf}
@@
 
-\begin{figure}[H]
-    \centering
+\begin{figure}
     \includegraphics{fig/prediction_modified_10p_increase_lags_5.pdf}
@@
 personal preferences. However, we are confident that our selected set of data captures the main aspects, as our model performs great on the provided test set.
+
+A repository containing the complete analyses and results can be found on \href{https://github.com/mxs01/AnalysingStudentDevelopment}{GitHub}.
 

### 2024-01-29 src/salaryExpHelpers.py

@@ -25,19 +25,20 @@ def plot(data: np.ndarray, forecast: np.ndarray, ogForecast: np.ndarray, years: 
     Returns:
             plt.Figure: creates matplotlib plot that shows difference between
     """
+    startPlottingAt = 12
     dataWithForecast = np.vstack((data, forecast))
     ogDataWithForecast = np.vstack((data, ogForecast))
-    yearsWithForecast = np.append(years, [f"{i+1}" for i in range(forecast.shape[0])])
+    yearsWithForecast = np.append(years, [f"prediction {i+1}" for i in range(forecast.shape[0])])
 
     plt.rcParams.update(bundles.icml2022(column="half", nrows=1, ncols=1, usetex=False))
     fig, ax1 = plt.subplots()
     plt.title('Enrolled students with altered salary')
     plt.xticks(rotation=30)
 
-    ax1.set_ylim([0, 10000])
+    ax1.set_ylim([0, 12000])
     ax1.set_ylabel('Total students', color=COL_STUDENT)
     # Plot main data
-    ax1.bar(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 0], color=COL_STUDENT, label='Enrolled students')
+    ax1.bar(yearsWithForecast[startPlottingAt:-len(forecast)], dataWithForecast[startPlottingAt:-len(forecast), 0], color=COL_STUDENT)
 
     # Plot forecast with a different color
     ax1.bar(yearsWithForecast[-len(forecast):], dataWithForecast[-len(forecast):, 0],
@@ -51,11 +52,11 @@ def plot(data: np.ndarray, forecast: np.ndarray, ogForecast: np.ndarray, years: 
     ax2.set_ylabel('Salary', color=COL_SALARY)
 
     # Plot main data
-    ax2.plot(yearsWithForecast[:-len(forecast)], dataWithForecast[:-len(forecast), 1], color=COL_SALARY, label='Average gross salary')
+    ax2.plot(yearsWithForecast[startPlottingAt:-len(forecast)], dataWithForecast[startPlottingAt:-len(forecast), 1], color=COL_SALARY, label='Average gross salary')
 
     # Plot forecast with a different color
     ax2.plot(yearsWithForecast[-len(forecast) - 1:], dataWithForecast[-len(forecast) - 1:, 1],
-             color=COL_SALARY_PRED, label='Average gross salary  forecast')
+             color=COL_SALARY_PRED, label='Average gross salary forecast')
 
[…]

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
increment

@@
 
-Our central contribution is a Vector autoregressive model (VAR) discussed in \Cref{sec:methods:var} that is built using just mentioned data. Combined with a
-partial autocorrelation function (PACF) the optimal value for the lag
-is determined, see \Cref{sec:methods:lag}, to ensure that relevant structure in the data is captured by the model.
+Our central contribution is a Vector autoregressive model (VAR) discussed in \Cref{sec:methods:var} that is built using just mentioned data.
+After processing (\Cref{sec:methods:processing}) the data, the model is optimized by different methods, see \Cref{sec:methods:lag}.
+This ensures that the relevant structure in the data is captured by the model.
 After verifying that the VAR model functions as expected and testing the model on a test set, \Cref{fig:accuracy model}, we
@@
 focused on these two instances.
-This allows us to conclude in \Cref{sec:methods:causality} that salary has a significant impact on the amount of STEM students. Further, we can model different salary development scenarios, 
-and see how the students would likely react (\Cref{sec:results}).
+This allows us to conclude in \Cref{sec:methods:causality} that salary expectations have a significant impact on the amount of STEM students.
+Further, we can model different salary development scenarios, and see how the students would likely react (\Cref{sec:results}).
 
@@
 \subsection{Choosing lag}\label{sec:methods:lag}
-To find the optimal hyperparameter we used two well-established methods, the PACF and information-based criteria as descriped in \citep{pacf}. Particularly the AIC and 
+To find the optimal hyperparameter we used two established methods, the PACF and information-based criteria as described in \citep{pacf}. Particularly the AIC and 
 BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ on $S_{n}$ to build a $(1-\alpha)$ confidence interval.
@@
 In our research, we used a vector autoregressive model to predict the amount of students in the next semester. 
-We used the salary data from the past 15 years and data about high school graduates in Baden-Würrtemberg as a prior. We optimized our Model to the best possible capacity,
+We used the salary data from the past 15 years and data about high school graduates in Baden-Würrtemberg as a prior. We optimized our Model to the best
+possible capacity,
 but there are strict limitations that can not be overcome. Vector autoregressive models are only capable of detecting linear patterns in the data.

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
Increment

@@
 The vector autoregressive model heavily depends on the stationarity assumption which ensures that the stochastical process stays the same over time. 
-In this context, this is assured by the following three axioms:
-\begin{samepage}
-\begin{enumerate}[label=\Roman*]
-    \item The mean $\mu$ is constant
-    \item The standard deviation is constant
-    \item There is no seasonality in the data
-\end{enumerate}    
-\end{samepage}
+In this context, this is assured by the following three axioms: The mean $\mu$ is constant, the standard deviation is constant
+and there is no seasonality in the data.
+
 Since we are only interested in predicting two time steps into the future, which is equivalent to one year,

### 2024-01-29 src/salaryExpHelpers.py
Increment

@@ -25,14 +25,14 @@ def plot(data: np.ndarray, forecast: np.ndarray, ogForecast: np.ndarray, years: 
     Returns:
             plt.Figure: creates matplotlib plot that shows difference between
     """
-    startPlottingAt = 12
+    startPlottingAt = 17
     dataWithForecast = np.vstack((data, forecast))
     ogDataWithForecast = np.vstack((data, ogForecast))
     yearsWithForecast = np.append(years, [f"prediction {i+1}" for i in range(forecast.shape[0])])
 
     plt.rcParams.update(bundles.icml2022(column="half", nrows=1, ncols=1, usetex=False))
     fig, ax1 = plt.subplots()
-    plt.title('Enrolled students with altered salary')
+    plt.title('Enrolled STEM students with altered salary')
     plt.xticks(rotation=30)
 
     ax1.set_ylim([0, 12000])
@@ -42,21 +42,22 @@ def plot(data: np.ndarray, forecast: np.ndarray, ogForecast: np.ndarray, years: 
 
     # Plot forecast with a different color
     ax1.bar(yearsWithForecast[-len(forecast):], dataWithForecast[-len(forecast):, 0],
-            color=COL_STUDENT_PRED, label='Enrolled students forecast')
+            color=COL_STUDENT_PRED, label='enrolled students forecast')
 
     # Plot original forecast with a different color
     ax1.bar(yearsWithForecast[-len(forecast):], ogDataWithForecast[-len(forecast):, 0],
-            color='silver', label='Original prediction', alpha=0.5)
+            color='silver', label='original prediction', alpha=0.5)
 
     ax2 = ax1.twinx()  # instantiate a second axes that shares the same x-axis
-    ax2.set_ylabel('Salary', color=COL_SALARY)
+    ax2.set_ylabel('Gross salary in EUR', color=COL_SALARY)
+#     ax2.set_ylabel(np.arange(3800, 4200, 100))
 
     # Plot main data
-    ax2.plot(yearsWithForecast[startPlottingAt:-len(forecast)], dataWithForecast[startPlottingAt:-len(forecast), 1], color=COL_SALARY, label='Average gross salary')
+    ax2.plot(yearsWithForecast[startPlottingAt:-len(forecast)], dataWithForecast[startPlottingAt:-len(forecast), 1], color=COL_SALARY, label='average gross salary')
 
     # Plot forecast with a different color
[…]

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
Increment
Increment [PDF]

@@
 With the obtained insights into the relation between the number of enrolled students, salary expectations, and number of graduates,
-it is now possible to make predictions on different scenarios.
+we are able to predict the upcoming four semesters, see \Cref{fig:predict}.
 
-The training of the model stays set, and the whole dataset available is used. It is now possible to 
-generate a new data set with a 10\% increase in salary expectations and set it as the new model input, \Cref{fig:altered salary}.
+\begin{figure}[H]
+    \includegraphics{fig/predictingStudents.pdf}
+    \caption{Predict the number of students enrolled in STEM courses.}
+    \label{fig:predict}
+\end{figure}
+
+It is also possible to construct different future scenarios.
+We generate a new data set with a 10\% increase in salary expectations, \Cref{fig:altered salary}.
+This will be used as a basis to predict from, while
+the unchanged data set is used for training.
 The observed change in the prediction supports the obtained results in \Cref{sec:methods:fit}. There is a significant
@@
 halfway through their academic careers.
-
 This also explains the huge spike in predicting the average gross salary. In the first few semesters,
 the amount of new students is not high enough, to capture the need for new people. Therefore the model suggests raising the
-salary even higher, to accommodate for that.
+salary even higher.
 
@@
 our prediction we had to replace our model with other machine learning applications.
-A promising approach for this could be to implement a Random Forest algorithm.\\
+A promising approach for this could be to implement a Random Forest algorithm.
 Our model on the other hand compensates for any potential lag in accuracy by being easily interpretable. Each parameter of the model influences exactly
@@
 personal preferences. However, we are confident that our selected set of data captures the main aspects, as our model performs great on the provided test set.
-
-A repository containing the complete analyses and results can be found on \href{https://github.com/mxs01/AnalysingStudentDevelopment}{GitHub}.
 
@@
 All authors jointly did the analyses of the data and wrote the text of the report.
+A repository containing the complete analyses and results can be found on \href{https://github.com/mxs01/AnalysingStudentDevelopment}{GitHub}.
+
 % Max Mustermann collected and prepared data.
@@
 
-% References
+% \newpage
 

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
Final version [PDF]

@@
 \begin{abstract}
-This paper investigates the role of expected salary in influencing the choice of academic majors among
-students in the STEM field at the University of Tübingen.
-The study focuses on the interplay among three key parameters: the enrollment of students, the expected salary, and the number of graduates.
-Employing a Vector Autoregressive model, we can confidently predict the number of students for the upcoming semesters
-and show that the expected salary plays a key role in making that prediction.
+
+Our research focus is to provide valuable insights
+into how future salaries affect the decision-making of students when it comes to selecting
+their major at the University of Tubingen. For this
+purpose we explore the dynamic relationship between student enrollment, expected salaries, and
+graduation rates. We are exploiting the Vector
+Autoregressive model to demonstrate a reliable
+forecasts for future STEM student numbers, underscoring the significant influence of projected
+income on these predictions.
 Furthermore, by manipulating the salary variable in our analysis, we observe a corresponding increase or decrease in the number of students.
@@
 
-Since we are only interested in predicting two time steps into the future, which is equivalent to one year,
+Since we are only interested in predicting two timesteps into the future, which is equivalent to one year,
 we can safely assume that our data is stationary.
@@
     \includegraphics{fig/params_Students.pdf}
-    \caption{Coefficients of VAR model to predict students time-series
+    \caption{Coefficients of VAR model to predict students time series
     for $t_0$. $t_{-1}$ to $t_{-5}$ are the values of each time series respectively at the lagged time step, which is equal to one semester.
@@
 % Workshops on Machine Learning (ML 1988 -- ML 1992). At the time this figure was
-% produced, the number of accepted papers for ICML 2008 was unknown and instead
+% produced, the number of accepted papers for ICML 2008 was unknown, and instead
 % estimated.}
@@
 
-% Max Mustermann collected and prepared data.
-Gabi Musterfrau and John Doe performed the data analysis.
-Jane Doe produced visualizations.
-All authors will jointly wrote the text of the report.
-Note that you, as a group, a collectively responsible for the report.
-Your contributions should be roughly equal in amount and difficulty.
-
-% \section*{Notes}
-
 % Your entire report has a \textbf{hard page limit of 4 pages} excluding references.
-(I.e.
-any pages beyond page 4 must only contain references).
-Appendices are \emph{not} possible.
-But you can put additional material, like interactive visualizations or videos, on a githunb repo (use \href{https://github.com/pnkraemer/tueplots}{links} in your pdf to refer to them).
-Each report has to contain \textbf{at least three plots or visualizations}, and \textbf{cite at least two references}.
-More details about how to prepare the report, inclucing how to produce plots, cite correctly, and how to ideally structure your github repo, will be discussed in the lecture, where a rubric for the evaluation will also be provided.
-
-% \newpage
+% (I.e. any pages beyond page 4 must only contain references). Appendices are \emph{not} possible.
+% But you can put additional material, like interactive visualizations or videos, on a githunb repo
+% (use \href{https://github.com/pnkraemer/tueplots}{links} in your pdf to refer to them).
+% Each report has to contain \textbf{at least three plots or visualizations}, and \textbf{cite at least two references}.
+% More details about how to prepare the report, inclucing how to produce plots, cite correctly, and how to ideally structure your github
+% repo, will be discussed in the lecture, where a rubric for the evaluation will also be provided.
 

### 2024-01-30 doc/DataLiteracyStudentProject/paper.tex
Final version [PDF]
Final Version [PDF]

@@
 \twocolumn[
-\icmltitle{A Prediction and Analysis of the interplay between Graduates and Salary Prospects in the STEM Field}
+\icmltitle{A Prediction and Analysis of the interplay between Students and Salary Prospects in the STEM Field}
 
@@
 that $S_{i,j}$ is also influenced by $S_{k,l}$ $i,k \in \{1,\dots, m\}: i\neq k$ and $j,l \in \{1,\dots, n-1\}: j\neq l$. As mentioned in \citep{var}
-A VAR model has one major advantage, over other possible approaches, which is its simplicity. Therefore its results remain interpretable.
+a VAR model has one major advantage, over other possible approaches, which is its simplicity. Therefore its results remain interpretable.
 This model only uses one hyperparameter which describes the amount of past data that is used to make a prediction. This parameter is called the lag.
@@
 \subsection{Choosing lag}\label{sec:methods:lag}
-To find the optimal hyperparameter we used two established methods, the PACF and information-based criteria as described in \citep{pacf}. Particularly the AIC and 
+To find the optimal hyperparameter we used two established methods, the partial autocorrelation function and information-based criteria as described in \citep{pacf}. Particularly the AIC and 
 BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ on $S_{n}$ to build a $(1-\alpha)$ confidence interval.
@@
 To get better results
-the AIC and BIC are used to find a better lag for the hyperparameter $p$, which is the lag. The abbreviation AIC
+the AIC and BIC are used to find a better lag for the hyperparameter $p$. The abbreviation AIC
 stands for Akaike Information Criterion and it tries to find an optimum between the complexity of the model and fitting the data.

### 2024-01-30 doc/DataLiteracyStudentProject/paper.tex
Final Version [PDF]

@@
 into how future salaries affect the decision-making of students when it comes to selecting
-their major at the University of Tubingen. For this
+their major at the University of Tübingen. For this
 purpose we explore the dynamic relationship between student enrollment, expected salaries, and
@@
 \section{Introduction}\label{sec:intro}
-In recent years, a growing number of high school graduates have chosen the path of higher education.
+In recent years, a growing number of high school graduates have chosen the path of higher education, according to \href{https://de.statista.com/infografik/27761/anzahl-eingeschriebener-mint-studentinnen-an-deutschen-universitaeten-nach-jahr/#:~:text=Technische%20Studiengänge&text=Rund%2037%20Prozent%20der%20drei,%2C%20Naturwissenschaft%20und%20Technik%2C%20eingeschrieben.}{Statista}.
 Science, technology, engineering, and mathematics, known as STEM, make up a large proportion of all students, although
